### 2014-07-22 Quality_Reports/Strand_exploration.Rmd
First commit cross - correlation per strand

## Exploratory analysis of the strand

### Description of the experiment

```{r package_options, include=FALSE}
opts_knit$set(progress = TRUE, verbose = TRUE)
```

![fig](figure/fromPaper/Kharchenko-Park-2008.jpg)