### 2015-10-27 rmarkdown/ChIPexo_paper.Rmd

---
title: "A quality control pipeline for ChIP-exo"
author: "Rene Welch"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
	html document:
		fig_caption: yes
		number_sections: yes
	md document:
		variant: markdown_github
	pdf document:
		fig_caption: yes
		number_sections: yes
		toc: yes
---

# Introduction

## What is ChIPexo ?

<!-- Here I am thinking to write the first section of the poster, which was -->
<!-- based on the Mendenhall and Bernstein note on Genome Biology -->

<!-- this paper need to include the analysis of the paper that dongjun made -->

### Literature review

#### ChIPexo related methods

<!-- About ChIPexo, there are: -->
<!-- CexoR -->
<!-- MACE -->
<!-- GEM (I think this two are the same) -->

#### ChIPseq methods used with ChIPexo

<!-- Some ChIPseq peak callers are used with ChIPexo regardless of they -->
<!-- work or not. Need to check which ones, if they work and if not perhaps -->
<!-- propose -->

### Biases present in ChIPexo data

#### Strand imbalance

#### Sequence digestion bias

### Reproducibility analysis

### Between protocols (ChIPseq vs ChIPexo)
### Between biological replicates

### Discussion

<!-- Here we are supposed to write the analysis that we made, the quality -->
<!-- control pipeline and how our claims were validated -->

### Conclusions

### Bibliography