### 2015-11-17 rmarkdown/chip_exo_paper.Rnw
Corrected orthography with ispell

%\VignetteEngine{knitr::knitr}
\documentclass{bmcart}

\usepackage{xcolor}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usepackage{float}
\usepackage{verbatim}
\usepackage{hyperref}
\hypersetup{
     colorlinks   = true,
     citecolor    = black
}
\hypersetup{linkcolor=black}
% \hypersetup{linkcolor=black}
% \hypersetup{linkcolor=black}

% \usepackage{authblk}
\usepackage[utf8]{inputenc} %unicode support

\begin{document}

%\maketitle

\begin{frontmatter}

\begin{fmbox}
\dochead{Draft}

\title{High Resolution Identification of Protein-DNA Binding
  Events and Quality Control for ChIP-exo data}

\author[
   addressref={aff6},                   % id's of addresses, e.g. {aff1,aff2}
   email={chungd@musc.edu}   % email address
]{\inits{DC}\fnm{Dongjun} \snm{Chung}}
\author[
   addressref={aff1},
   email={welch@stat.wisc.edu}
]{\inits{RW}\fnm{Rene} \snm{Welch}}
\author[
   addressref={aff3},
   email={ong@cs.wisc.edu}
]{\inits{IO}\fnm{Irene} \snm{Ong}}
\author[
   addressref={aff3,aff4},
   email={jagrass@wisc.edu}
]{\inits{JG}\fnm{Jeffrey} \snm{Grass}}
\author[
   addressref={aff3,aff4,aff5},
   email={landick@bact.wisc.edu}
]{\inits{RL}\fnm{Robert} \snm{Landick}}
\author[
   addressref={aff1,aff2},
   corref={aff1},
   email={keles@stat.wisc.edu}
]{\inits{SK}\fnm{S\"und\"uz} \snm{Kele\c{s}}}

\address[id=aff1]{%                           % unique id
  \orgname{Department of Statistics, University of Wisconsin Madison}, % university, etc
  \street{1300 University Avenue},                     %
  %\postcode{}                                % post or zip code
  \city{Madison},                              % city
  \cny{WI}                                    % country
}

\address[id=aff2]{%
  \orgname{Department of Biostatistics and Medical Informatics, University of Wisconsin Madison},
  \street{600 Highland Avenue},
%  \postcode{24105}
  \city{Madison},
  \cny{WI}
}
\address[id=aff3]{
  \orgname{Great Lakes Bioenergy Research Center, University of Wisconsin Madison},
  \street{1552 University Avenue},
  \city{Madison},
  \cny{WI}
}
\address[id=aff4]{
  \orgname{Department of Biochemistry, University of Wisconsin Madison},
  \street{433 Babcock Drive},
  \city{Madison},
  \cny{WI}
}
\address[id=aff5]{
  \orgname{Department of Bacteriology, University of Wisconsin Madison},
  \street{1550 Linden Drive},
  \city{Madison},
  \cny{WI}
}
\address[id=aff6]{
  \orgname{Department of Public Health Sciences, Medical University of South Carolina},
  \street{135 Cannon Street},
  \city{Charleston},
  \cny{SC}
}

% \begin{artnotes}
% %\note{Sample of title note}     % note to the article
% \note[id=n1]{Equal contributor} % note, connected to author
% \end{artnotes}

\end{fmbox}

\begin{abstractbox}

  \begin{abstract}

  Recently, ChIP-exo has been developed to investigate protein-DNA
  interaction in higher resolution compared to popularly used
  ChIP-Seq. Although ChIP-exo has drawn much attention and is
  considered as powerful assay, currently, no systematic studies have
  yet been conducted to determine optimal strategies for experimental
  design and analysis of ChIP-exo. In order to address these
  questions, we evaluated diverse aspects of ChIP-exo and found the
  following characteristics of ChIP-exo data. First, Background of
  ChIP-exo data is quite different from that of ChIP-Seq data. However,
  sequence biases inherently present in ChIP-Seq data still exist in
  ChIP-exo data. Second, in ChIP-exo data, reads are located around
  binding sites much more tightly and hence, it has potential for high
  resolution identification of protein-DNA interaction sites, hence
  the space to allocate the reads is greatly reduced. Third, although
  often assumed in the ChIP-exo data analysis methods, the peak pair
  assumption does not hold well in real ChIP-exo data. Fourth, spatial
  resolution of ChIP-exo is comparable to that of PET ChIP-Seq and
  both of them are significantly better than resolution of SET
  ChIP-Seq. Finally, for given fixed sequencing depth, ChIP-exo
  provides higher sensitivity, specificity, and spatial resolution
  than PET ChIP-Seq.

  In this article, we provide a quality control pipeline which
  visually asses ChIP-exo biases and calculates a signal-to-noise
  measure. Also, we updated dPeak \cite{dpeak}, which makes a striking
  balance in sensitivity, specificity, and spatial resolution for
  ChIP-exo data analysis.

\end{abstract}
\begin{keyword}
  \kwd{ChIP-exo}
  \kwd{QC}
  \kwd{TFBS}
  \kwd{BS identification on High-Res}
\end{keyword}

\end{abstractbox}

\end{frontmatter}

\newpage

* to remove later

\tableofcontents

\newpage

\listoffigures

\newpage

\section{Background}
\label{sec:intro}

ChIP-exo (Chromatin Immunopecipitation followed by exonuclease
digestion and next generation sequencing) Rhee and Pugh (\cite{exo1})
is the state-of-the-art experiment developed to attain single
base-pair resolution of protein binding site identification and it is
considered as a powerful alternative to popularly used ChIP-Seq
(Chromatin Immunoprecipitation coupled with next generation sequencing
) assay. ChIP-exo experiments first capture millions of DNA fragments
(150 - 250 bp in length) that the protein under study interacts with
using random fragmentation of DNA and a protein-specific
antibody. Then, exonuclease is introduced to trim 5' end of each DNA
fragment to a fixed distance from the bound protein. As a result,
boundaries around the protein of interest constructed with 5' ends of
fragments are located much closed to bound protein compared to
ChIP-Seq. This is the step unique to ChIP-exo that could potentially
provide significantly higher spatial resolution compared to
ChIP-Seq. Finally, high throughput sequencing of a small region (25 to
100 bp) at 5' end of each fragment generates millions of reads or
tags.

While the number of produced ChIP-exo data keeps increasing,
characteristics of ChIP-exo data and optimal strategies for
experimental design and analysis of ChIP-exo data are not fully
investigated yet, including issues of sequence biases inherent to
ChIP-exo data, choice of optimal statistical methods, and
determination of optimal sequencing depth. However, currently, the
number of available ChIP-exo data is still limited and their
sequencing depths are still insufficient for such investigation. To
address this limitation we gathered ChIP-exo data from diverse
organisms: CTCF factor in human \cite{exo1}; ER factor in human and
FoxA1 factor in mouse from \cite{exoillumina}; and generated
$\sigma^{70}$ factor in Escherichia coli (E. Coli) under aerobic ($ +
O_2$) condition, and treated by rifampicin by 0 and 20 minutes.

DNA libraries generated by the ChIP-exo protocol seem to be less
complex than the libraries generated by ChIP-Seq
\cite{exo_review}. Hence, most of current guidelines \cite{encode_qc}
may not be applicable on ChIP-exo, additionally to our knowledge there
are not established quality control pipelines for ChIP-exo. To address
this challenge, we suggest a series of quality control visualizations
to understand which biases are present in ChIP-exo data. Related to
quality control, Previous ChIP-exo analysis used ChIP-Seq samples to
compare the resolution between experiments (\cite{exo1}, \cite{exo2},
\cite{exoillumina}). In \cite{carroll.qc}, Carroll et. al. studied the
use of the Strand-Cross Correlation (SCC) \cite{strandcc}. and showed
that by filtering blacklisted regions the estimation of the SCC is
improved. However, this method requires to know blacklisted regions in
advance which may not be available. In our pipeline we propose two
out-the-shelf metrics equivalent to RSC and NSC for the estimation of
the signal-to-noise ratio for a ChIP-exo sample.

In order to archive the potential benefits of ChIP-exo on protein
binding site identification, it is critical to understand which are
the important characteristics of ChIP-exo data and to use algorithms
that could fully utilize information available in ChIP-exo data. Rhee
and Pugh \cite{exo1} discussed that reads in the forward and reverse
strand might construct peak pairs around bound protein, of which
heights were implicitly assumed to be symmetric. Hence, they used the
``peak pair method'' that predicts the midpoint of two modes of peak
pairs as potential binding site. Mace \cite{mace}, CexoR \cite{cexor}
and peakzilla \cite{peakzilla}, recently developed ChIP-exo data
analysis methods, are also based on this peak pair
assumption. However, appropriateness of such assumption was not fully
evaluated in the literature yet.  Furthermore, it is still unknown
which factors could affect protein binding site identification using
ChIP-exo data. In order to address this problem, we investigated
various aspects of ChIP-exo data by contrasting them with their
respective ChIP-Seq experiments.

Currently, research on statistical methods for ChIP-exo data is still
in its very early stage. Although many methods have been proposed to
identify protein binding sites from ChIP-Seq data (reviewed in
\cite{evaluation} and \cite{computation}), such as MACS \cite{macs},
CisGenome \cite{cisgenome} and MOSAiCS \cite{mosaics}, these
approaches reveal protein binding sites in lower resolution, i.e., at
an interval of hundreds to thousands of base pairs. Furthermore, they
report only one ``mode'' or ``predicted binding location'' per
peak. Hence, these methods are not appropriate to evaluate the
potential of ChIP-exo data for high resolution identification of
protein binding sites. More recently, deconvolution algorithms such as
Deconvolution \cite{csdeconv}, GEM \cite{gem} (an improved version of
\cite{gps}) and PICS \cite{pics} have been proposed to identify
binding sites in higher resolution using ChIP-Seq data. However, most
of them are still not tailored for ChIP-exo and PET and SET ChIP-Seq
data in a unified framework and as a result, currently available
methods are not appropriate for fair comparison between ChIP-exo and
ChIP-Seq. To address these limitations, we developed an improved of
dPeak \cite{dpeak}, a high resolution binding site identification
(deconvolution) algorithm that we previously developed for PET and SET
ChIP-Seq data, so that it can also handle ChIP-exo data. The dPeak
algorithm implements a probabilistic model that accurately describes
the ChIP-exo and ChIP-Seq data generation process.

In this paper, we demonstrate that the peak pair assumption of Rhee
and Pugh \cite{exo1} does not hold well in real ChIP-exo
data. Furthermore, we found that when we analyze ChIP-exo data from
eukaryotic genomes, it is important to consider sequence biases
inherent to ChIP-exo data, such as mappability and GC content in order
to improve sensitivity and specificity of binding site
identification. dPeak outperforms or performs competitively with
ChIP-exo data analysis such as GEM and MACE. More importantly, when
comparable number of reads is used for both ChIP-exo and ChIP-Seq,
dPeak couple with ChIP-exo data provides resolution comparable to PET
ChIP-Seq and both significantly improve the resolution of protein
binding site identification compared to SET-based analysis with any of
the available methods.

\newpage

\section{Results and discussion}
\label{sec:results}

\subsection{Deeply sequenced E.Coli $\sigma^{70}$ ChIP-exo and ChIP-Seq data}
\label{sec:ourdata}

$\sigma^{70}$ factor is a transcription initiation factor of
housekeeping genesin E. coli. In E. coli genomes, many promoters
contain multiple transcription start sites (TSS) and these TSS are
often closely space (10 ~ 150 bp). These closely spaced binding sites
are considered to be multiple ``switches'' that differentially
regulate gene expression under diverse growth conditions
\cite{regdb_old}, \cite{regulondb}. Therefore, investigation of
ChIP-exo's potential for identification and differentiation of closely
spaced binding sites are invaluable for elucidating the
transcriptional networks of prokaryotic genomes.

\subsection{Current guidelines and quality metrics}
\label{sec:qcnow}

To have a first view on the data, we calculated a series of the most
commonly used quality indicators for ChIP-Seq data, which are listed
in table \ref{tab:qcbase}. The PBC, NSC and RSC were calculated as
defined in \cite{encode_qc} and the SSD was calculated as defined in
\cite{htseq}. Additionally, we calculated the FSR to compare the
number of fragments in each strand. A few conclusions that we may get
by observing the PBC in table \ref{tab:qcbase} are all the samples
have a quite low library complexity, hence it would be suggested to
repeat the experiment.

\begin{table}[h]
  \centering
\begin{tabular}{l|r|r|r|r|r|r}
\hline\hline
Condition & Rep. & Nr. reads & PBC & SSD & FSR & NSC  \\
\hline\hline
Aerobic & 1 & 1,454,566 & 0.2156 & 0.0830 & 0.5012 & 114.14 \\
\hline
Aerobic & 2 & 864,714 & 0.2471 & 0.0488 & 0.4923 & 160.93  \\
\hline
Rif-0min & 1 & 960,256 & 0.2823 & 0.0361 & 0.4998 & 10.29 \\
\hline
Rif-0min & 2 & 2,247,295 & 0.2656 & 0.1091 & 0.5047 &  25.08  \\
\hline
Rif-20min & 1 & 1,940,387 & 0.2698 & 0.0820 & 0.5054 & 17.69  \\
\hline
Rif-20min & 2 & 4,229,574 & 0.2153 & 0.1647 & 0.5071 &  14.11 \\
\hline
\end{tabular}  
  \caption{Usual quality control indicators applied to our
    $\sigma^{70}$ samples. PBC stands PCR-bottleneck coefficient, SSD
    for standardized standard deviation, FSR for forward strand ratio,
    NSC for normalized strand cross-correlation coefficient. We
    omitted the RSC (relative strand cross-correlation since ChIP-exo
    experiments don't usually have an accompanying input file)}

  \label{tab:qcbase}
\end{table}

\begin{figure}[h!]
  \centering
  \includegraphics[width = .8\textwidth]{../figs/for_paper/EColi_strand_cross_corr.pdf}
  \caption{Strand cross-correlation curves for $\sigma^{70}$
    samples. The ``phantom peak'' and the summit that corresponds to
    the read and fragment length respectively are confounded, due to
    the exonuclease enzyme digestion.}
  \label{figsup:scc}
\end{figure}

In a good ChIP-Seq experiment the strand cross-correlation is
maximized at the fragment length and there is also a ``phantom peak''
at the sample's read length (\cite{encode_qc} and \cite{strandcc}).
However, in ChIP-exo's case this two peaks are confounded. Hence an
enrichment measure based in the strand cross-correlation such as NSC
or RSC is harder to interpret. In figure \ref{figsup:scc} we plotted
the SCC curves for $\sigma^{70}$ ChIP-exo experiments used to
calculate the NSC and RSC values in table \ref{tab:qcbase}. According
to ENCODE's quality metrics, NSC values close to one corresponds to
low enrichment samples or bad quality datasets, hence in our case the
last four samples would be considered to have higher quality than they
actually have. On the other hand, all PBC values are considerably low
and correspond to samples with severe bottlenecking issues, in
particular we are going to show that for the first two samples we are
obtaining a smaller resolution and identifying a comparable number of
targets as in ChIP-Seq PET experiments.

\subsection{ChIP-exo Quality Control pipeline}
\label{sec:QC}

Figure \ref{fig:qcdiagram} shows a flowchart for the ChIPexoQC
pipeline. Which basically partitions genome by keeping the
non-digested ChIP-exo regions. Then, for each region it calculates a
series of independent statistics. Finally, it creates several
visualizations designed to asses the overall quality levels of A
ChIP-exo sample.

\begin{figure}[h!]
  \centering
  \includegraphics[width = .9\textwidth]{../figs/for_paper/coverage_diagram.pdf}
  \caption{Diagram of the ChIP-exo Quality control pipeline(this is a
    place-holder). The genome is partitioned into several regions by
    removing the empty segments and for each regions we calculate a
    collection of qc indicators.}
  \label{fig:qcdiagram}
\end{figure}

\subsubsection{Enrichment and library complexity in ChIP-exo data}
\label{sec:enri}

In ChIP-exo experiments, the background is often digested by the
exonuclease enzyme, therefore to determine the sample's quality is
necessary to address the balance between the enrichment and library
complexity. To diagnose this, we considered the Average Read Coverage
(ARC) and the Unique Read Coverage Rate (URCR) which are defined as:

\begin{align}
  \mbox{ARC} &= \frac{\text{Nr. of reads in the region}}{\text{Width of the region}} \nonumber \\
  \mbox{URCR} &= \frac{\text{Nr. of reads mapped to only one position
      in the region}}{\text{Nr. of reads in the region}} \nonumber
\end{align}

Figure \ref{fig:enrich}A shows hexbin plots illustrating the
interaction between these two quantities. Notice how there are two
strong arms in each panel: The first one corresponds to regions with
low $\mbox{ARC}$ values and $\mbox{URCR}$ values that can vary across
the whole $[0,1]$ interval, while the second one shows a decreasing
trend in $\mbox{URCR}$ as $\mbox{ARC}$ increases. When an experiment
shows a higher degree of enrichment, then the separation of this two
arms is more noticeable, since the second arm corresponds to possibly
enriched regions (\ref{fig:enrich2}A and \ref{fig:enrich2}B).

<<strata_param , include =FALSE,echo = FALSE, eval = TRUE>>=

  nstrata <- 100

@

Figure \ref{fig:enrich}B shows boxplots of the $\mbox{local-NSC}$
stratified by the number of reads mapped exactly to one position for
all $\sigma^{70}$ ChIP-exo experiments; for each stratum, we sampled
\Sexpr{nstrata} regions that whenever possible (in the opposite case,
all the regions in the category were considered). There is an
increasing trend as the number of unique position increases, which
means that this is an effective indicator of library complexity since
regions where the reads are mapped to more positions are usually more
complex.

%% perhaps do two other slots about how this works... a good experiment vs a bad experiment
%% of the same region

\begin{figure}[h!]
  \centering
  \includegraphics[width = .8\textwidth]{../figs/for_paper/Sig70_enrichment.pdf}
  \includegraphics[width = .8\textwidth,page =3]{../figs/for_paper/Local_SCC_indicator_by_strata.pdf}
  \caption{A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$ and
    B) boxplots of the $\mbox{local-NSC}$ stratified by nr. of reads
    mapped to only one position in the regions for the ChIP-exo
    $\sigma^{70}$ experiments divided by condition and biological
    replicate}
  \label{fig:enrich}
\end{figure}

\subsubsection{Strand imbalance in ChIP-exo data}
\label{sec:strand_imbalance}

The strand imbalance assessment is based in the observation that the
enriched regions usually have a higher concentration of fragments,
therefore we examine the FSR (defined as the ratio of the number of
forward stranded reads divided by the total number of reads in a given
region) as the region with lower depth are being filtered out. This
indicator is of particular importance, since several methods rely on
the ``peak-pair'' assumption. In table \ref{tab:qcbase}, we calculated
the FSR and noticed that for all the samples, it's value is close to
0.5, which means that there are roughly the same amount of reads in
both strands. However, figure \ref{fig:comp}B shows that this
assumption don't hold in practice.

In order to asses the strand imbalance we create the following
visualization in figure \ref{fig:strand}. A) presents the FSR's
behavior as lower depth regions are being filtered out. while B)
shows which percentage of the regions are composed by fragments of
both strand or only one. Roughly speaking in a lower quality dataset
there is going to be a higher amount of single stranded-regions. To
asses this, we divided the ChIP-exo regions into two sets, the ones
that overlap with a known set of peaks and the opposite and performed
$\chi^2$ tests.

\begin{figure}[h!]
  \centering  
  \includegraphics[width = .8\textwidth]{../figs/fsr/draft_fsr_plot.pdf}
  \includegraphics[width = .8\textwidth]{../figs/fsr/draft_label_prop_plot.pdf}
  \caption{ABCD}
  \label{fig:strand}
\end{figure}

\subsection{Comparison with ChIP-Seq data}
\label{sec:comp}

<<comp_param,include = FALSE,echo = FALSE,eval = TRUE>>=

  bin_size <- 150
  frag_len <- 150

@

We first compared various factors that could affect binding site
identification between ChIP-exo and ChIP-Seq data. in order to compare
distribution of signal and background between ChIP-exo and ChIP-Seq
data, we calculated ChIP tag counts across the genome by counting the
number of reads mapping to each of \Sexpr{bin_size} non-overlapping
window after extending reads by \Sexpr{frag_len} to their 3' end
directions. ChIP tag counts in ChIP-exo data were linearly related to
ChIP tag counts in ChIP-Seq data for the regions with high ChIP tag
counts (Figure \ref{fig:comp}A). This implies that signals for
potential binding sites are well reproducible between ChIP-exo and
ChIP-Seq data. On the other hand, there was clear difference in the
background distribution between them. In ChIP-Seq data reads were
almost uniformly distributed over background (non-binding) regions and
the ChIP tag counts in there regions were significantly larger than
zero. In contrast, in ChIP-exo data, there was larger variation in
ChIP tag counts among background regions and ChIP tag counts were much
lower in these regions compared to ChIP-Seq data. There were also
large proportion of regions without any read in ChIP-exo data. These
results indicate that background distribution of ChIP-exo data is less
homogeneous than that of ChIP-Seq data.

We next evaluated the ``peak pair'' assumption of Rhee and Pugh
\cite{exo1}, MACE \cite{mace}, CexoR, \cite{cexor} and peakzilla
\cite{peakzilla}, i.e. a peak of reads in the forward strand is
usually paired with a peak of reads in the reverse strand that is
located in the other site of the binding site. In order to evaluate
this assumption, we reviewed the proportion of reads in the forward
strand in candidate regions (i.e. regions with at least one binding
site) in $\sigma^{70}$ ChIP-exo data. We found that strands of reads
were much less balanced in ChIP-exo data than in ChIP-Seq data in
these regions with potential binding sites (Fig. \ref{fig:comp}B) and
this indicates that the peak pair assumption might not hold in real
ChIP-exo data.

%  We obtained esentially the same conclusion from the
% ChIP-exo data for CTCF factor from human sample Rhee and Pugh
% (\cite{exo1}) as well.

\begin{figure}[h]
  \centering
  \includegraphics[width = .46\textwidth,page = 3 ]{../figs/for_paper/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
  \includegraphics[width = .46\textwidth]{../figs/for_paper/forward_strand_ratio_comp_old.pdf}
  \caption{ A) ChIP signal in ChIP-exo was linearly related to that in
    PET ChIP-Seq data in the region with high ChIP tag counts. In
    contrast, there were clear differences in their background
    distribution, where several ChIP-exo background regions were
    empty. B) In ChIP-exo data, strands of reads were significantly
    less balanced in the regions with potential binding sites compared
    to SET ChIP-Seq data. }
  \label{fig:comp}
\end{figure}

\subsubsection{Comparison with ChIP-Seq data using dPeak}
\label{sec:dpeak_analysis}

<<reso_param, include = FALSE , eval  = TRUE , echo = FALSE>>=

  ext <- 20

@

Figure \ref{fig:reso_all} shows different comparisons among ChIP-exo,
PET ChIP-Seq and SET ChIP-Seq. A regulonDB annotation was considered
identified if the distance between it and dPeak binding site estimate
was at most of \Sexpr{ext} bp. That way, the sensitivity is defined as
the proportion of regulonDB annotations identified in a peak. In
figure \ref{fig:reso_all}A can be seen that the sensitivity of all
protocols increases as the average distance between binding sites
does. Despite that when the binding events in a peak are closer to
each other both ChIP-exo and PET ChIP-Seq are comparable, as the
distance increases ChIP-exo identifies a higher proportion of the
regulonDB annotations; additionally SET ChIP-Seq is significantly less
sensible than both ChIP-exo and PET ChIP-Seq. In figure
\ref{fig:reso_all}B the distance between a regulonDB annotation to its
closest prediction is compared for ChIP-exo, PET and SET ChIP-Seq;
while the first two are comparable, both outperform SET
ChIP-Seq.

In figures \ref{fig:reso_all}C and \ref{fig:reso_all}D, we observe the
behavior of dPeak's estimated parameters for single end data (both
ChIP-exo and ChIP-Seq). The $\delta$ parameter measures the distance
from the 5' end of the fragment to its respective binding event and
the $\sigma$ parameter indicate the fragments distribution
variability around their respective binding sites (more information in
\cite{dpeak}). For both parameters, it can be seen that the ChIP-exo
estimates are smaller than the SET ChIP-Seq estimates in average,
which agrees with the fact that the sample reads are allocated more
tightly around the binding events in ChIP-exo data. Hence we can
conclude that the peaks shape is quite different between ChIP-exo and
SET ChIP-Seq.

\begin{figure}[h!]
  \centering
  \includegraphics[width = .46\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
  \includegraphics[width = .46\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
   \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/sigma_delta_old_densities.pdf}
   \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/sigma_delta_old_densities.pdf}

   \caption{Comparison of (A) sensitivity and (B) resolution between
     ChIP-exo and ChIP-Seq data. Sensitivity is defined as the
     proportion of regulonDB annotations identified using each
     data. Resolution is defined as the distance between regulonDB
     annotation and its closest prediction. $\quad$ C) $\delta$ parameter in
     dPeak measures average distance of the reads to their respective
     binding site. In ChIP-exo data, reads were located much closer to
     the binding site than in SET ChIP-Seq. D) $\sigma$ parameter
     measure the dispersion of reads around each binding site. In
     ChIP-exo data, reads showed less variation around the their
     respective binding sites compared to SET ChIP-Seq.}
  \label{fig:reso_all}
\end{figure}

\subsection{Recommendations for the design of ChIP-exo experiments}
\label{sec:reco}

<<design_param , include = FALSE , echo = FALSE , eval = TRUE>>=

  K <- 10
  ext <- 15

@

We sampled a fixed among of fragments for each of the ChIP-exo, PET
ChIP-Seq and SET ChIP-Seq datasets of the $\sigma^{70}$ sample in
aerobic conditions. For each sampled dataset we applied our
lower-to-higher resolution pipeline by calling peaks with MOSAiCS
\cite{mosaics} and then deconvolving the binding events by using dPeak
\cite{dpeak}. For the ChIP-exo datasets we called peaks by using
GC-content and mappability with MOSAiCS, and for the ChIP-Seq datasets
we used their respective Input samples.

\begin{figure}[h!]
  \centering
%  \includegraphics[width = .8\textwidth]{../figs/for_paper/Sig70_aerobic_saturation.pdf}
  \includegraphics[width = .8\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
  \caption{Comparison of the number of candidate regions (A),
    predicted events (B), identified targets (C) and resolution (D)
    among ChIP-exo, PET ChIP-Seq and SET ChIP-Seq. RegulonDB
    annotations are considered as a gold standard. A gold standard was
    considered identified if a BS was estimated inside a \Sexpr{ext}
    bp window around it}
  \label{fig:design}
\end{figure}

Figure \ref{fig:design} shows the behavior of each data type when
their depth is fixed. It is remarkable that even when the number of
candidate peaks or the number of predicted events in quite lower than
for both ChIP-Seq cases, it outperforms them in the number of
identified targets and in resolution. This may suggest that with
ChIP-exo less false positive peaks are being called and that when the
targets are being identified, dPeak is able to estimate binding
locations closer to the actual true cases. Additionally, we can see
that as the read depth increases all four indicators do so as well,
which may indicate that with ChIP-exo a smaller amount of reads is
necessary to identify a higher number of targets, but it may also be
possible that this is an artifact occurring due to ChIP-exo's lower
library complexity.

\section{Conclusions}
\label{sec:conclusions}

\newpage

\section{Methods}
\label{sec:methods}

\color{red}

\subsection{Growth conditions}
\label{sec:growth}

\subsection{ChIP-exo experiments}
\label{sec:experiments}

\subsection{Library preparation, sequencing and mapping of sequencing reads}
\label{sec:library}

\color{black}

\subsection{Method comparison with ChIP-exo}
\label{sec:suppcomp}

We considered dPeak Chung et al. \cite{dpeak}, GEM Guo et
al. \cite{gem} and MACE Wang et al. \cite{mace} for ChIP-exo data
analysis. For the dPeak algorithm, we used the R package \emph{dPeak}
version 1.0.0, which is available from
\url{http://dongjunchung.github.io/dpeak/}. For the GEM algorithm we
used its Java implementation version 0.9, which is available from
\url{http://cgs.csail.mit.edu/gem/}. For the MACE algorithm, we used
its Python implementation version 1.0, which is available from
\url{http://dldcc-web.brc.bcm.edu/lilab/MACE/docs/html/}. For the
PEAKZILLA algorithm, we used its Python implementation, which is
available in \url{http://www.starklab.org/data/peakzilla/}. Candidate
regions for dPeak were identified for each replicate of ChIP-exo data
using the MOSAiCS algorithm Kuan et al. \cite{mosaics} (one sample
analysis using false discovery rate of 0.0001) implemented as an R
package \emph{mosaics} 2.4 (available from \emph{Bioconductor}
\url{https://www.bioconductor.org/packages/release/bioc/html/mosaics.html}). We
further filtered out candidate region with average ChIP tag count less
than 3,000 to avoid potential false positives based in exploratory
analysis. These regions were also explicitly provided to the GEM
algorithm as candidate regions. Default tuning parameters were used
during model fitting for all methods.

% And for
% the CexoR algorithm, we used its R implementation version 1.8.0 which
% is available from
% \url{http://bioconductor.org/packages/release/bioc/html/CexoR.html}

\subsection{Local-NSC}
\label{sec:localnsc}

The local strand cross-correlation is defined as:

\begin{align}
  \mbox{local-NSC} &= \frac{max_{x_\delta} f(x_\delta)}{\sigma} \nonumber \\
 y_\delta &= f(x_\delta) + \sigma \epsilon_\delta \nonumber
\end{align}

where $f$ is calculated by fitting a local polynomial regression model
over the strand cross-correlation calculated using exclusively the
fragments that overlap a pre-defined genomic region, $y_\delta$ is the
correlation of the forward and backward coverages when one is shifted
$x_\delta$ bp towards their 3' end.

The local polynomial model was fitted using the \emph{loess} function
from the R software version 3.2.1 using the default tuning parameters.

%% not sure if here or before it would be nice to add an example of how this 
%% method works. probably here

\newpage

\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{chip_exo_paper}

\nocite{exo_gb}
\nocite{maplot1}
\nocite{maplot2}
% \nocite{esl}

\newpage

\section*{Supplement}
\label{sec:supp}

\subsection*{Additional Enrichment plots for $\sigma^{70}$}
\label{sec:enrichsup}

In figure \ref{fig:enrich2} it is shown the same relationship as in
figure \ref{fig:enrich}, but considering only regions formed where the
reads are being allocated in more than 10 (A) and 30 (B) unique
positions respectively. Both plots show that the vertical arms formed
by regions with low $\mbox{ARC}$ is formed by low complexity regions,
that way suggesting that this segment correspond to the background of
a ChIP-exo experiment.

\begin{figure}[h!]
  \centering
  \includegraphics[width = .8\textwidth,page = 3]{../figs/for_paper/Sig70_enrichment.pdf}
  \includegraphics[width = .8\textwidth,page = 4]{../figs/for_paper/Sig70_enrichment.pdf}
  \caption{Hexbin plots of $\mbox{ARC}$ vs $\mbox{URCR}$ for each
    region after partitioning the genome. In A and B the regions with
    reads mapped to at most 10 and 30 positions respectively where not
    considered. $\mbox{ARC}$ is defined as the ratio of the nr. of
    reads and the width of a region and $\mbox{URCR}$ is the ratio of
    the number of unique position where the reads are being allocated
    and the number of reads in a region.}
  \label{fig:enrich2}
\end{figure}

\end{document}