### 2015-11-24 rmarkdown/prelims_paper.Rnw
made some minor corrections in the paper, and fixed the figures to be named as reps instead of the bam files

\documentclass[11pt]{article}
%\documentclass[11pt]{scrartcl}
\usepackage{amsmath, amsfonts, amssymb}
\usepackage{adjustbox}
\usepackage{fullpage}
\usepackage{epsfig}
\usepackage{hyperref}
\hypersetup{
     colorlinks   = true,
     citecolor    = black
}
\hypersetup{linkcolor=black}

\usepackage{multicol}

\renewcommand{\baselinestretch}{1.4}
\usepackage{float}
\usepackage{wrapfig}

\usepackage{tikz}
\usetikzlibrary{shapes,arrows}

\title{High Resolution Identification of Protein-DNA Binding Events
  and Quality Control for ChIP-exo data\vspace*{\fill}}

\author{Rene Welch\\Preliminary Examination\\Department of Statistics,
  University of Wisconsin-Madison}

\date{December 1st, 2015}

%% commmands
%\usepackage{Sweave}

\begin{document}

\newcommand{\sig}{\sigma^{70}}

\maketitle

\vspace*{\fill}

\textbf{Committee Members:}

\textbf{Professor S\"und\"uz Kele\c{s}}, Department of Statistics,
Department of Biostatistics and Medical Informatics

\textbf{Professor Karl Broman}, Department of Biostatistics and
Medical Informatics

\textbf{Professor Colin Dewey}, Department of Computer Sciences,
Department of Biostatistics and Medical Informatics

\textbf{Professor Christina Kendziorski}, Department of Biostatistics
and Medical Informatics

\textbf{Professor Ming Yuan}, Department of Statistics

\thispagestyle{empty}

\newpage

\tableofcontents

\newpage

\listoffigures

\newpage

\section*{Abstract}

    Recently, ChIP-exo has been developed to investigate protein-DNA
    interaction in higher resolution compared to popularly used
    ChIP-Seq. Although ChIP-exo has drawn much attention and is
    considered as powerful assay, currently no systematic studies
    have yet been conducted to determine optimal strategies for
    experimental design and analysis of ChIP-exo. In order to address
    these questions, we evaluated diverse aspects of ChIP-exo and
    found the following characteristics of ChIP-exo data. First, the
    background of ChIP-exo data is quite different from that of
    ChIP-Seq data. However, sequence biases inherently present in
    ChIP-Seq data still exist in ChIP-exo data. Second, in ChIP-exo
    data, reads are located around binding sites much more tightly and
    hence, it has potential for high resolution identification of
    protein-DNA interaction sites, and also the space to allocate the
    reads is greatly reduced. Third, although often assumed in the
    ChIP-exo data analysis methods, the ``peak pair'' assumption does
    not hold well in real ChIP-exo data. Fourth, spatial resolution of
    ChIP-exo is comparable to that of PET ChIP-Seq and both of them
    are significantly better than resolution of SET ChIP-Seq. Finally,
    for given fixed sequencing depth, ChIP-exo provides higher
    sensitivity, specificity and spatial resolution than PET
    ChIP-Seq.

    We provide a quality control pipeline which visually assess
    ChIP-exo biases, library complexity and enrichment; and calculates
    a signal-to-noise measure. Also, we updated dPeak (Chung et al.,
    2012 \cite{dpeak}), which makes a striking balance in sensitivity,
    specificity and spatial resolution for ChIP-exo data analysis.

\newpage

\section{Introduction}
\label{sec:intro}

ChIP-exo (Chromatin Immunopecipitation followed by exonuclease
digestion and next generation sequencing) Rhee and Pugh, 2011
(\cite{exo1}) is the state-of-the-art experiment developed to attain
single base-pair resolution of protein binding site identification and
it is considered as a powerful alternative to popularly used ChIP-Seq
(Chromatin Immunoprecipitation coupled with next generation sequencing
 assay.

While the number of produced ChIP-exo data keeps increasing,
characteristics of ChIP-exo data and optimal strategies for
experimental design and analysis of ChIP-exo data are not fully
investigated yet, including issues of sequence biases inherent to
ChIP-exo data, choice of optimal statistical methods, and
determination of optimal sequencing depth. However, currently the
number of available ChIP-exo data is still limited and their
sequencing depths are still insufficient for such investigation. To
address this limitation we gathered ChIP-exo data from diverse
organisms: CTCF factor in human \cite{exo1}; ER factor in human and
FoxA1 factor in mouse (Serandour et al., 2013 \cite{exoillumina}); and
generated $\sig$ factor in Escherichia Coli (E. Coli) under
aerobic ($ + O_2$) condition, and treated by rifampicin by 0 and 20
minutes.

DNA libraries generated by the ChIP-exo protocol seem to be less
complex than the libraries generated by ChIP-Seq (Mahony et al., 2015
\cite{exo_review}). Hence, most of current QC guidelines (Landt et
al., 2012 \cite{encode_qc}) may not be applicable on ChIP-exo,
additionally to our knowledge there are not established quality
control pipelines for ChIP-exo. To address this challenge, we suggest
a collection of quality control visualizations to understand which
biases are present in ChIP-exo data. Previous ChIP-exo analysis used
ChIP-Seq samples to compare the resolution between experiments
(\cite{exo1}, \cite{exo2}, \cite{exoillumina}); Carroll et al., 2014
\cite{carroll.qc} studied the use of the Strand-Cross Correlation
(SCC) (Kharchenko et al., 2008 \cite{strandcc}) and showed that by
filtering blacklisted regions the estimation of the SCC is
improved. However, this method requires to know blacklisted regions in
advance which may not be available, additionally using SCC may not be
useful since the peaks that are attained at the read and fragment
length are confused in a typical ChIP-exo SCC curve.  In our pipeline
we propose two out-the-shelf analysis: an enrichment plot and the
local normalized SCC cofficient.

In order to archive the potential benefits of ChIP-exo on protein
binding site identification, it is critical to understand which are
the important characteristics of ChIP-exo data and to use algorithms
that could fully utilize information available in ChIP-exo data. Rhee
and Pugh, 2011 \cite{exo1} discussed that reads in the forward and
reverse strand might construct peak pairs around bound protein, of
which heights were implicitly assumed to be symmetric. Hence, they
used the ``peak pair method'' that predicts the midpoint of two modes
of peak pairs as potential binding site. Mace (Wang et al., 2014
\cite{mace}), CexoR (Madrigal, 2015 \cite{cexor}) and peakzilla
(Bardet et al., 2013 \cite{peakzilla}), recently developed ChIP-exo
data analysis methods, are also based on this peak pair
assumption. However, appropriateness of such assumption was not fully
evaluated in the literature yet.  Furthermore, it is still unknown
which factors could affect protein binding site identification using
ChIP-exo data. In order to address this problem, we investigated
various aspects of ChIP-exo data by contrasting them with their
respective ChIP-Seq experiments.

Currently, research on statistical methods for ChIP-exo data is still
in its very early stage. Although many methods have been proposed to
identify protein binding sites from ChIP-Seq data (reviewed by
Wilbanks and Facciotti, 2012 \cite{evaluation} and Pepke and Wold,
2009 \cite{computation}), such as MACS (Zhang et al., 2008
\cite{macs}), CisGenome (Hongkai et al., 2008 \cite{cisgenome}) and
MOSAiCS (Kuan et al., 2009 \cite{mosaics}), these approaches reveal
protein binding sites in lower resolution, i.e., at an interval of
hundreds to thousands of base pairs. Furthermore, they report only one
``mode'' or ``predicted binding location'' per peak. Hence, these
methods are not appropriate to evaluate the potential of ChIP-exo data
for high resolution identification of protein binding sites. More
recently, deconvolution algorithms such as Deconvolution (Lun et al.,
2009 \cite{csdeconv}), GEM (Guo et al., 2012 \cite{gem}, an improved
version of Guo et al., 2010 \cite{gps} ) and PICS (Zhang et al., 2010
\cite{pics}) have been proposed to identify binding sites in higher
resolution using ChIP-Seq data. However, most of them are still not
tailored for ChIP-exo and PET and SET ChIP-Seq data in a unified
framework and as a result, currently available methods are not
appropriate for fair comparison between ChIP-exo and ChIP-Seq. To
address these limitations, we developed an improved of dPeak (Chung et
al., 2013 \cite{dpeak}), a high resolution binding site identification
(deconvolution) algorithm that we previously developed for PET and SET
ChIP-Seq data, so that it can also handle ChIP-exo data. The dPeak
algorithm implements a probabilistic model that accurately describes
the ChIP-exo and ChIP-Seq data generation process.

In this work, we demonstrate that the ``peak pair'' assumption of
Rhee and Pugh \cite{exo1} does not hold well in real ChIP-exo
data. Furthermore, we found that when we analyze ChIP-exo data from
eukaryotic genomes, it is important to consider sequence biases
inherent to ChIP-exo data, such as mappability and GC content in order
to improve sensitivity and specificity of binding site
identification. We evaluated several method to identify binding events
and dPeak outperforms or performs competitively respecto to GEM and
MACE when analyzing ChIP-exo data. More importantly, when comparable
number of reads is used for both ChIP-exo and ChIP-Seq, dPeak coupled
with ChIP-exo data provides resolution comparable to PET ChIP-Seq and
both significantly improve the resolution of protein binding site
identification compared to SET-based analysis with any of the
available methods.

\section{ChIP-exo and ChIP-Seq review}
\label{sec:exo}

\begin{wrapfigure}{l}{.4\textwidth}
\centering
\includegraphics{../figs/diagrams/ChIPexo_diagram.png}
\caption{ChIP-exo diagram. The only difference with ChIP-Seq is the
  second step, where an exonuclease enzyme is added and it digest the
  DNA fragments starting from the 5' end until it finds a crosslink
  protein (this figure is taken from Furey, 2012 \cite{chipbeyond}).}
\label{fig:exo}
\end{wrapfigure}

ChIP-exo is based in a modification to the ChIP-seq protocol, where an
exonuclease enzyme is added and it digests the fragments starting by
the 5' ends and stops until it finds a protein that is crosslinked to
the DNA.

ChIP-exo experiments first capture millions of DNA fragments
(150 - 250 bp in length) that the protein under study interacts with
using random fragmentation of DNA and a protein-specific
antibody. Then, exonuclease is introduced to trim 5' end of each DNA
fragment to a fixed distance from the bound protein. As a result,
boundaries around the protein of interest constructed with 5' ends of
fragments are located much closed to bound protein compared to
ChIP-Seq. This step is unique to ChIP-exo and could potentially
provide significantly higher spatial resolution compared to
ChIP-Seq. Finally, high throughput sequencing of a small region (25 to
100 bp) at 5' end of each fragment generates millions of reads or
tags.

Since this protocol is a modification to ChIP-Seq, some of its
characteristics are still maintained in ChIP-exo while several new are
being found. Rhee and Pugh, 2011 \cite{exo1}, showed that ChIP-exo
generated fragments are located around the binding events more tightly
than ChIP-Seq generated fragments. Hence, it is of utmost importance
to understand how this new protocol works, specifically we focused on
understanding which biases are mantained from ChIP-Seq to ChIP-exo,
which biases are specific to ChIP-exo, how the current ChIP-Seq QC
guidelines behave in ChIP-exo and we developed a new QC pipeline
specific to ChIP-exo. A typical example is shown in figure
\ref{fig:exo_example}, which shows the $\sig$ and $\beta'_f$ factors
(and the $\beta$ factor only on top) on a E. Coli genome's region:
First, both regions show that the $\sig$ sample is enriched; second
the $\sig$ sample in the ChIP-exo panel shows that its respective form
in the ChIP-Seq panel may be form of more binding events; and third,
either of the $\beta$ factors show that the complexity of the ChIP-exo
libraries may be lower in some cases, i.e. the peak is mapped with
fewer positions.

\begin{figure}[H]
  \centering
  \includegraphics[width = .8 \textwidth,page = 107]{/p/keles/ChIPexo/volume5/OpenClosedComplexes/figs/peaks/ChIPseq_ChIPexo_peaks_rif20.pdf}
  \caption{An example of a PET ChIP-Seq peak compared to
    ChIP-exo. $\sig$, $\beta$ and $\beta_f'$ are subunits of
    the transcription initiation complex in E. Coli. Gene regions on
    the reverse strand are highlighted in light blue and the region
    was centered respect to ChIP-Seq $\sig$ summit. The library
    complexity is lower in ChIP-exo than in ChIP-Seq but the
    resolution is increased, this can be observed at the $\sig$
    peak that for ChIP-exo was created as the combination of at least
    three ChIP-exo ``sub-peaks''.}
  \label{fig:exo_example}
\end{figure}

\subsection{Current Quality Control Measures for ChIP-Seq}
\label{sec:QC_chipseq}

To our knowledge it doesn't exists a QC pipeline for ChIP-exo. Hence,
we analyzed how the current ChIP-Seq QC measures behave on ChIP-exo
data.

Table \ref{tab:qcbase} contains measures that are commonly used with
ChIP-Seq data. PBC stands for PCR bottleneck amplification and is
defined as the ratio between the number of positions to which exactly
one mappping read maps divided by the number of position to which at
least one mapping reads maps, the ENCODE project proposed a series of
ranges to interpret this quantity (which can be found in
\url{https://www.encodeproject.org/data-standards/2012-quality-metrics/}),
but it doesn't apply to ChIP-exo, if it did all the samples with
$\mbox{PBC} < 0.5$ should be reject due to be showing severe
bottlenecking. Planet et al., 2011 proposed the Standarized Standard
Deviation which defined as:

\[
\mbox{SSD} = \mbox{SD}_N / \sqrt{N} 
\]

where $\mbox{SD}_N$ is the standard deviation of the coverage of the
sample and $N$ is the number of reads. This measure is smaller for
ChIP-exo since the reads are located more tightly around binding
sites, but it may be biased since several fragments are being
digested by exonuclease enzyme, hence there are several position in
the genome where no fragment is being mapped. Finally the normalized
strand cross-correlation is defined as:

\[
\mbox{NSC} = \frac{ \max_\delta y(\delta)}{\min_\delta y(\delta)}
\]

where $y(\delta)$ is strand cross-correlation defined as in (\ref{scc}).

\subsubsection{Strand cross-correlation}
\label{sec:scc}

The strand cross-correlation (SCC) was proposed by Kharchenko et al.,
2008 \cite{strandcc} and it may be one of the most used of the
ChIP-Seq QC metrics. The SCC curve is defined as:

\begin{align}
  y(\delta) = \sum_c w_c r\left[ n_c^+ \left(x + \frac{\delta}{2}
    \right), n_c^- \left( x- \frac{\delta}{2} \right)\right]
\label{scc}
\end{align}

where $y(\delta)$ is the SCC for a strand shift $\delta$, $r$ is the
Pearson correlation, $w_c$ is the propotion of tags mapped to
chromosome $c$ and $n_c^S$ is the tag count vector for strand $S$ and
chromsome $c$.

For ChIP-Seq data, the SCC will have two peaks: One when $\delta$ is
equal to reads length and one when $\delta$ is equal to the unobserved
fragment length, ideally the SCC function is maximized at this
point. Landt et al., 2012 \cite{encode_qc} explain that a succesful if
this happens, it is marginally succesful if its maximized at the read
length peak but still shows the fragment length summit and failed if
is maximized at the read length peak but the second summit is not
present.

\begin{figure}[H]
  \centering  
\includegraphics[width = .7\textwidth]{../figs/for_paper/scc_ctcf.pdf}
  \caption{SCC for CTCF factor in HeLa cell line for ChIP-exo (obtained from \cite{exo1}) and SET ChIP-Seq (obtained from \cite{encode1}) }
  \label{fig:scc}
\end{figure}

Figure \ref{fig:scc} shows the SCC function for both a ChIP-exo (in
red) and a SET ChIP-Seq (in green) for CTCF in HeLa cell line. The
ChIP-Seq dataset behaves as the succesful case from \cite{encode_qc},
but the ChIP-exo doesn't. Additionally, since the exonuclease enzyme
digests the DNA fragments, then the shifts where the SCC may reach a
locall or a global maximum are confounded. Hence, new QC measures are
necessary for ChIP-exo data.

\begin{table}[H]
  \centering
  \caption{Usual quality control indicators applied to the $\sigma^{70}$ samples. PBC stands PCR-bottleneck coefficient, SSD for standardized standard deviation, NSC for normalized strand cross-correlation coefficient (Another measure related to the strand cross-correlation curve is the RSC was ommited since input sample usually don't exists for ChIP-exo). $\sig$ samples were given by Robert Landick, FoxA1 and ER samples are from Serandour et al., 2013 \cite{exoillumina} and CTCF sample is from Rhee and Pugh, 2011 \cite{exo1}.}

\begin{tabular}{l|l|r|r|r|r|r|r}
\hline\hline
IP & Organism  &Condition & Rep. & Nr. reads & PBC & SSD &  NSC  \\
\hline\hline
$\sig$ & E.Coli & Rif-0min & 1 & 960,256 & 0.2823 & 0.0361 &  10.29 \\
\hline
$\sig$ &  E.Coli & Rif-0min & 2 & 2,247,295 & 0.2656 & 0.1091  &  25.08  \\
\hline
$\sig$ &  E.Coli & Rif-20min & 1 & 1,940,387 & 0.2698 & 0.0820  & 17.69  \\
\hline
$\sig$ &  E.Coli & Rif-20min & 2 & 4,229,574 & 0.2153 & 0.1647  &  14.11 \\
\hline
FoxA1 &  Mouse &  NA & 1 & 22,210,461 & 0.6562 & 9.12 $\times 10^{-5}$ & 21.452 \\
\hline
FoxA1 &  Mouse &  NA & 2 & 22,307,557 & 0.7996 & 7.94 $\times 10^{-5}$ & 60.661 \\
\hline
FoxA1 &  Mouse &  NA & 3 & 22,421,729 & 0.1068 & 1.31 $\times 10^{-4}$ & 72.312 \\
\hline
ER & Human & NA & 1 & 9,289,835 & 0.8082 & 3.64 $\times 10^{-5}$ & 19.843 \\
\hline
ER & Human & NA & 2 & 11,041,833 & 0.8024 & 4.6 $\times 10^{-5}$ & 21.422 \\
\hline
ER & Human & NA & 3 & 12,464,836 & 0.8203 & 4.89 $\times 10^{-5}$ & 19.699 \\
\hline
CTCF & Human & NA & 1 &   48,478,450 & 0.4579 & 1.29 $\times 10^{-4}$ & 15.977 \\
\hline
\end{tabular}  
  \label{tab:qcbase}
\end{table}

\vspace*{\fill}

\subsection{The dPeak model}
\label{sec:dpeak}

Chung et al., 2013 proposed the dpeak model to identifity
transcription factor binding sites in high resolution. Two models are
proposed: One to deconvolve the signal of single end tags (SET) data,
i.e. when only one end of the DNA fragment is sequenced; and another
to deconvolve the signal of paired end tags (PET) data, i.e. when both
ends of the DNA fragment are sequenced. For ChIP-exo, both ends are
sequenced but only the 5' end is considered, since is the side where
the exonuclease enzyme digests the DNA.

Given a peak region with $m$ positions, with $n$ DNA fragments and that
the region is generated by a known $g^*$ amount of binding
events. Each fragment is described by the start position where the
starting end is aligned and the strand from which the fragment is
sequenced. Since only one end is sequenced, then the fragment length
is not observed. dPeak assumes that the strand $D_i$ follows a
Bernoulli distribution with known parameter $p_D$, and the start
position $R_i$ for the $i$-th fragment is generated according to the
following procedure:

\begin{itemize}
\item If the fragment was sequenced from the forward strand ($D_i =
  1$):
  \begin{itemize}
  \item The reads belongs to the background: $R_i | Z_i = 0, D_i = 1
    \sim \mbox{Unif}(1 - \beta + 1, m)$
  \item The read belong to the $g$-th binding event: $R_i | Z_i = g,
    D_i = 1 \sim \mbox{N}(\mu_g - \delta , \sigma^2)$
  \end{itemize}
\item If the fragment was sequenced from the backward strand ($D_i =
  0$):
  \begin{itemize}
  \item The reads belongs to the background: $R_i | Z_i = 0, D_i = 1
    \sim \mbox{Unif}(1 , m + \beta  - 1)$
  \item The read belong to the $g$-th binding event: $R_i | Z_i = g,
    D_i = 1 \sim \mbox{N}(\mu_g + \delta , \sigma^2)$
  \end{itemize}

\end{itemize}

and the binding event to which the $i$-th fragment belongs $Z_i = g$
is generated from a

$\mbox{Multinomial}(\pi_0,\pi_1,\cdots,\pi_{g^*})$.

\subsubsection{dPeak's algorithm initialization strategy}
\label{sec:algo}

dPeak's model likelihood is optimized by the stochastic EM
algorithm. Hence, it is of utmost importance how the parameters are
initialized in order to obtain a global maximum.

Chung et al., 2013 \cite{dpeak} derived an initialization strategy for
the stochastic EM algorithm where each region is initialized by
separate independently and the binding sites locations $\mu_g,
g=1,\cdots,g^*$ are initialized by a uniform partition.

For this work, we considered a new strategy where the all the regions
are ranked by a statistical measure (as the $-\log(p)$ where $p$ is
the empirical p.value of the test $H_0: \mathcal{R} \text{ is
  enriched}$, where $\mathcal{R}$ is a genomic region) or a empirical
quantity (as the number of ChIP tag counts in the region) and use as
initial estimates of the $\delta$ and $\sigma$ parameters the average
of the estimates from the highest ranked $N_{\text{top}}$ peaks.

\subsection{Comparison with ChIP-Seq data}
\label{sec:comp_chipseq}

<<comp_param,include = FALSE,echo = FALSE,eval = TRUE>>=

  bin_size <- 150
  frag_len <- 150

@

We first compared various factors that could affect binding site
identification between ChIP-exo and ChIP-Seq data. In order to compare
distribution of signal and background between ChIP-exo and ChIP-Seq
data, we calculated ChIP tag counts across the genome by counting the
number of reads mapping to each of \Sexpr{bin_size} non-overlapping
window after extending reads by \Sexpr{frag_len} to their 3' end
directions. ChIP tag counts in ChIP-exo data were linearly related to
ChIP tag counts in ChIP-Seq data for the regions with high ChIP tag
counts (Figure \ref{fig:comp}A). This implies that signals for
potential binding sites are well reproducible between ChIP-exo and
ChIP-Seq data. On the other hand, there was clear difference in the
background distribution between them. In ChIP-Seq data reads were
almost uniformly distributed over background (non-binding) regions and
the ChIP tag counts in there regions were significantly larger than
zero. In contrast, in ChIP-exo data, there was larger variation in
ChIP tag counts among background regions and ChIP tag counts were much
lower in these regions compared to ChIP-Seq data. There were also
large proportion of regions without any read in ChIP-exo data. These
results indicate that background distribution of ChIP-exo data is less
homogeneous than that of ChIP-Seq data.

\begin{figure}[H]
  \centering
  \includegraphics[width = .46\textwidth,page = 3 ]{../figs/for_paper/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
  \includegraphics[width = .46\textwidth]{../figs/for_paper/forward_strand_ratio_comp_old.pdf}
\includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
  \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
  \caption{ A) ChIP signal in ChIP-exo was linearly related to that in
    PET ChIP-Seq data in the region with high ChIP tag counts. In
    contrast, there were clear differences in their background
    distribution, where several ChIP-exo background regions were
    empty. B) In ChIP-exo data, strands of reads were significantly
    less balanced in the regions with potential binding sites compared
    to SET ChIP-Seq data. C) ChIP tag counts increase linearly as
    mappability scores increases. D) ChIP tag counts increase linearly
    as GC content score increases when GC content is less than 0.6 and
    then ChIP tag counts decrease as GC content increases. }
  \label{fig:comp}
\end{figure}

We next evaluated the ``peak pair'' assumption from Rhee and Pugh,
2011 \cite{exo1}, i.e. a peak of reads in the forward strand is
usually paired with a peak of reads in the reverse strand that is
located in the other site of the binding site. Wang et al., 2014
\cite{mace}, Madrigal 2015 \cite{cexor} and Bardet et al., 2013
\cite{peakzilla} proposed method rely in this assumption. In order to
evaluate this assumption, we reviewed the proportion of reads in the
forward strand in candidate regions (i.e. regions with at least one
binding site) in $\sig$ ChIP-exo data. We found that strands of reads
were much less balanced in ChIP-exo data than in ChIP-Seq data in
these regions with potential binding sites (Fig. \ref{fig:comp}B) and
this indicates that the peak pair assumption might not hold in real
ChIP-exo data.

We evaluated ChIP-exo data for CTCF factor from human genome
\cite{exo1} to investigate issues specific to eukaryotic genomes for
binding sites identification. Figures \ref{fig:comp}C and
\ref{fig:comp}D display the bin-level average read counts against
mappability and GC content. Each data point is obtained by averaging
the read counts across bins with the same mappability of GC
content. These results indicate that binding site identification in
ChIP-exo sample might also benefit from the use of methods that take
into account of apparent sequence biases such as mappability and GC
content.

\section{Results}
\label{sec:results}

\subsection{ChIP-exo Quality Control pipeline}
\label{sec:QC}

Figure \ref{fig:qcdiagram} shows a flowchart for the ChIPexoQC
pipeline. Which basically partitions the genome by keeping the
non-digested ChIP-exo regions. Then, for each region it calculates a
series of statistics. Finally, it creates several visualizations
designed to assess the quality levels of a ChIP-exo sample.

\begin{figure}[H]
  \centering
  \includegraphics[width = .7\textwidth]{../figs/for_paper/coverage_diagram2.pdf}
  \caption{Diagram of the ChIP-exo Quality control pipeline. The
    genome is partitioned into several regions by removing the empty
    segments and for each regions we calculate a collection of QC
    indicators, then for each region $R_k$ a vector of summary
    statistics $T(R_k)$ is calculated.}
  \label{fig:qcdiagram}
\end{figure}

\subsubsection{Local Normalized Strand cross-correlation}
\label{sec:locNSC}

We considered the creation of a novel QC measure for ChIP-exo called
local normalized SCC or $\mbox{local-NSC}$. This measure

\begin{align}
  y(\delta) = r\left[ n^+ \left(x + \frac{\delta}{2}
    \right), n^- \left( x- \frac{\delta}{2} \right)\right]
\nonumber
\end{align}

where again $r$ is the Pearson correlation but $n^+$ and $n^-$ are the
coverage vectors of a given region in the genome. Figure
\ref{fig:locSCC} shows that as the library complexity of the region is
lower, then the local SCC seems to show higher variance. Hence, we
fit a nonparametric regression model to smooth the SCC signal:

\begin{align}
 y(\delta) &= f(x_\delta) + \epsilon_\delta \nonumber
\end{align}

where $y(\delta)$ is the observed SCC for shift $x_\delta$,
$\epsilon_\delta \sim N(0,\sigma_f^2)$. Finally the $\mbox{local-NSC}$
is defined as a signal-to-noise ratio by using:

\begin{align}
    \mbox{local-NSC} &= \frac{max_{x_\delta} \hat{f}(x_\delta)}{\hat{\sigma}_f} \label{locNSC}
\end{align}

\begin{figure}[H]
  \centering
  \includegraphics[width = .8\textwidth]{../figs/for_paper/local_SCC_example.pdf}
  \caption{Using the FoxA1 sample from \cite{exoillumina}. A) shows an
    example region for the 3 replicates. The library complexity seem
    to be comparable between the two top rows and both outperfom the
    last row. The number of reads for the first replicate is higher
    than the number of reads for the other replicates. B) Shows the
    local SCC for the 3 replicates, this measure is comparable among
    the 3 replicates. The first replicate shows the lowest residual
    error, while the third one shows the highest. In blue, a ``loess''
    regression model is fitted.}
\label{fig:locSCC}
\end{figure}

\subsubsection{Enrichment and library complexity in ChIP-exo data}
\label{sec:enri}

In ChIP-exo experiments, the background is often digested by the
exonuclease enzyme, therefore to determine the sample's quality is
necessary to address the balance between the enrichment and library
complexity. To diagnose this, we consider the Average Read Coverage
(ARC) and the Unique Read Coverage Rate (URCR) which are defined as:

\begin{align}
  \mbox{ARC} &= \frac{\text{Nr. of reads in the region}}{\text{Width of the region}} \nonumber \\
  \mbox{URCR} &= \frac{\text{Nr. of reads mapped to only one position
      in the region}}{\text{Nr. of reads in the region}} \nonumber
\end{align}

Using the mouse-FoxA1 experiment from \cite{exoillumina} and the
relationship between this two quantities, library complexity and
sample enrichment was explored. Figure \ref{fig:enrich}A shows hexbin
plots illustrating the interaction between these two
quantities. Notice how there are two strong arms in each panel: The
first one corresponds to regions with low $\mbox{ARC}$ values and
varying $\mbox{URCR}$ values across the $(0,1]$ interval, while the
second one shows a decreasing trend in $\mbox{URCR}$ as $\mbox{ARC}$
increases. When an experiment shows a higher degree of enrichment,
then the separation of this two arms is more noticeable, since the
second arm corresponds to possibly enriched regions
(\ref{fig:enrich2}A and \ref{fig:enrich2}B).

<<strata_para , include =FALSE,echo = FALSE, veal = TRUE>>=

  strata <- 400
  high <- 100
  med <- 50
  low <- 20

@

Figure \ref{fig:enrich}B shows box plots of the $\mbox{local-NSC}$ on
regions of the three replicates, stratified by the number of reads
mapped exactly to one position for all $\sigma^{70}$ ChIP-exo
experiments; The high stratum is defined by regions consisting of more
than \Sexpr{high} unique positions, the medium stratum for regions
where the number of unique position is on the
(\Sexpr{med},\Sexpr{high}) range and the low stratum in the
(\Sexpr{low},\Sexpr{med}) range. For each stratum and biological
replicate, we sampled \Sexpr{strata} regions whenever possible (in the
opposite case, all the regions in the category were considered) and
calculated the $\mbox{local-NSC}$ for those regions. For all three
replicates, it is shown an increasing trend as the number of unique
position increases, which means that the $\mbox{local-NSC}$ is an
effective indicator of library complexity. Figure \ref{fig:enrich}C
shows the $\mbox{local-NSC}$ coefficient calculated for the same set
of regions, stratified by the number of unique positions in the
regions of the first replicate using the same criteria as before. In
the high panel it is shown that the first replicate sample performs
slightly better than the second and both outperform the third one,
this agrees with \ref{fig:enrich}A, where only in the left-most panel
doesn't seem to be a separation between the low $\mbox{ARC}$ arm and
the other.

\begin{figure}[H]
  \centering
  \includegraphics[width = .7\textwidth]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
  \includegraphics[width = .7\textwidth,page =3]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata.pdf}
%  \includegraphics[width = .7\textwidth,page =4]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata_sameRegion.pdf}
  \caption{Using the mouse-FoxA1 experiment from \cite{exoillumina}:
    A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$, in general
    we can see a slight separation into two strong arms, one
    corresponds to low $\mbox{ARC}$ and varying $\mbox{URCR}$, and for
    the other $\mbox{URCR}$ decreases as $\mbox{ARC}$ increases B)
    Boxplots of the $\mbox{local-NSC}$ stratified by nr. of reads
    mapped to only one position in the regions for the ChIP-exo
    experiments divided by condition and biological replicate.}
  \label{fig:enrich}
\end{figure}

% and C) Boxplots of the a set of the first replicate regions,
% stratified as in (B) with the $\mbox{local-NSC}$ calculated with the
% different biological replicates.

\subsubsection{Strand imbalance in ChIP-exo data}
\label{sec:strand_imbalance}

The strand imbalance assessment is based in the observation that the
enriched regions usually have a higher concentration of fragments,
therefore we examined the FSR (defined as the ratio of the number of
forward stranded reads divided by the total number of reads in a given
region) as the region with lower depth are being filtered out. This
indicator is of particular importance, since several methods rely on
the ``peak-pair'' assumption. In table \ref{tab:qcbase}, we calculated
the FSR and noticed that for all the samples, it's value is close to
0.5, which means that there are roughly the same amount of reads in
both strands. However, figure \ref{fig:comp}B shows that this value is
not representative of the sample locally, therefore the assumption
doesn't hold in practice.

<<fsr,include = FALSE,eval = TRUE,echo = FALSE>>=

  de <- 100

  pval <- list()
  pval[[1]] <- 0.1039
  pval[[2]] <- 0.0064
  pval[[3]] <- 0.0871

@

In order to assess the strand imbalance we created the following
visualization presented in figure \ref{fig:strand}. A) shows the FSR's
behavior as lower depth regions are being filtered out, while B) shows
which percentage of the regions are composed by fragments of both
strands or only one (forward or backward). As the number of reads per
region's lower bound increases, the quantiles tend to approach the
median. To evaluate this statement, for each replicate of the FoxA1
dataset, we divided the regions formed by at least \Sexpr{de}
fragments into into the ones that overlap with any of the sample's
peaks. Then we considered a Kolmogorov-Smirnov tests to assess if the
distribution of both groups were the same. The p.value for the third
replicate's test is \Sexpr{pval[[1]]}, for the first replicate's test
is \Sexpr{pval[[2]]} and for the second test is
\Sexpr{pval[[3]]}. This shows that only the middle sample from figure
\ref{fig:strand} is the least imbalance sample from the the three,
while the other two still show a higher imbalance at the \Sexpr{de}
level.

\begin{figure}[H]
  \centering  
  \includegraphics[width = .8\textwidth,page = 3]{../figs/Carroll_mice_for_paper/Strand_imbalance.pdf} 
  \caption{Strand imbalance QC plots for the same data as in figure
    \ref{fig:enrich}. A) FSR distribution quantiles as the lower depth
    regions are being filtered out, all quantiles approach to the
    median as the lower bound increases. B) Stacked histogram with the
    proportion of regions that are formed by two strands or only one,
    in a good sample the single-stranded regions are going to be
    filtered out quickly as in the middle row.}
  \label{fig:strand}
\end{figure}

\subsection{Comparison with ChIP-Seq data using dPeak}
\label{sec:dpeak_analysis}

<<reso_param, include = FALSE , eval  = TRUE , echo = FALSE>>=

  ext <- 20

@

Figure \ref{fig:reso_all} shows different comparisons among ChIP-exo,
PET ChIP-Seq and SET ChIP-Seq. A RegulonDB annotation was considered
identified if the distance between it and dPeak binding site estimate
was at most of \Sexpr{ext} bp. That way, the sensitivity is defined as
the proportion of RegulonDB annotations identified in a peak. In
figure \ref{fig:reso_all}A can be seen that the sensitivity of all
protocols increases as the average distance between binding sites
does. Despite that when the binding events in a peak are closer to
each other, both ChIP-exo and PET ChIP-Seq are comparable, as the
distance increases ChIP-exo identifies a higher proportion of the
RegulonDB annotations; additionally SET ChIP-Seq is significantly less
sensible than both ChIP-exo and PET ChIP-Seq. In figure
\ref{fig:reso_all}B the distance between a RegulonDB annotation to its
closest prediction is compared for ChIP-exo, PET and SET ChIP-Seq;
while the first two are comparable, both outperform SET ChIP-Seq.

In figures \ref{fig:reso_all}C and \ref{fig:reso_all}D, we observe the
behavior of dPeak's estimated parameters for single end data (both
ChIP-exo and ChIP-Seq). The $\delta$ parameter measures the distance
from the 5' end of the fragment to its respective binding event and
the $\sigma$ parameter indicates the fragments distribution
variability around their respective binding sites (more information in
\cite{dpeak}). For both parameters, it can be seen that the ChIP-exo
estimates are smaller than the SET ChIP-Seq estimates in average,
which agrees with the fact that the sample reads are allocated more
tightly around the binding events in ChIP-exo data. Hence we can
conclude that the peaks shape is quite different between ChIP-exo and
SET ChIP-Seq.

\begin{figure}[H]
  \centering
  \includegraphics[width = .46\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
  \includegraphics[width = .46\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
\end{figure}

\begin{figure}[H]
\centering
   \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/sigma_delta_old_densities.pdf}
   \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/sigma_delta_old_densities.pdf}

   \caption{Comparison of (A) sensitivity and (B) resolution between
     ChIP-exo and ChIP-Seq data. Sensitivity is defined as the
     proportion of RegulonDB annotations identified using each
     data. Resolution is defined as the distance between RegulonDB
     annotation and its closest prediction. $\quad$ C) $\delta$ parameter in
     dPeak measures average distance of the reads to their respective
     binding site. In ChIP-exo data, reads were located much closer to
     the binding site than in SET ChIP-Seq. D) $\sigma$ parameter
     measure the dispersion of reads around each binding site. In
     ChIP-exo data, reads showed less variation around the their
     respective binding sites compared to SET ChIP-Seq.}
  \label{fig:reso_all}
\end{figure}

\subsection{Recommendations for the design of ChIP-exo experiments}
\label{sec:reco}

<<design_param , include = FALSE , echo = FALSE , eval = TRUE>>=

  K <- 10
  ext <- 15

@

We sampled a fixed amount of fragments for each of the ChIP-exo, PET
ChIP-Seq and SET ChIP-Seq datasets of the $\sigma^{70}$ sample in
aerobic conditions. For each sampled dataset we applied our
lower-to-higher resolution pipeline by calling peaks with MOSAiCS
\cite{mosaics} and then deconvolving the binding events by using dPeak
\cite{dpeak}. For the ChIP-exo datasets we called peaks by using
GC-content and mappability with MOSAiCS, and for the ChIP-Seq datasets
we used their respective Input samples.

\begin{figure}[H]
  \centering
%  \includegraphics[width = .8\textwidth]{../figs/for_paper/Sig70_aerobic_saturation.pdf}
  \includegraphics[width = .8\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
  \caption{Comparison of the number of candidate regions (A),
    predicted events (B), identified targets (C) and resolution (D)
    among ChIP-exo, PET ChIP-Seq and SET ChIP-Seq. RegulonDB
    annotations are considered as a gold standard. A gold standard was
    considered identified if a BS was estimated inside a \Sexpr{ext}
    bp window around it}
  \label{fig:design}
\end{figure}

Figure \ref{fig:design} shows the behavior of each data type when
their depth is fixed. It is remarkable that even when the number of
candidate peaks or the number of predicted events is lower for
ChIP-exo, it outperforms both PET and SET ChIP-Seq in the number of
identified targets and resolution. This may suggest that with ChIP-exo
less false positive peaks are being called and that when the targets
are being identified, dPeak estimates binding locations closer to the
true location. Additionally, we can see that as the read depth
increases all four indicators do so as well, which may indicate that
with ChIP-exo a smaller amount of reads is necessary to identify a
higher number of targets, but it may also be possible that this is an
artifact occurring due to ChIP-exo's lower library complexity.

\section{Software packages}
\label{sec:software}

For this project several R packages were created or updated:

\begin{itemize}
\item \textbf{dPeak}: We updated the initialization strategy. The
  latest version is currently available from
  \url{http://dongjunchung.github.io/dpeak/}.
\item \textbf{ChIPexoQual}: This package contains the QC pipeline for
  ChIP-exo. The last version is available in
  \url{https://github.com/welch16/ChIPexoQual}.
\item \textbf{Segvis}: The goal of this package is to visualize
  genomic regions by using aligned reads such as in figures
  \ref{fig:exo} or \ref{fig:qcdiagram}.  The latest version is
  available in \url{https://github.com/keleslab/Segvis}.
\item \textbf{ChIPUtils}: This package attempts to gather the most
  commonly used ChIP-Seq QC (as in table \ref{tab:qcbase} or figure
  \ref{fig:scc}) indicators and additional utilities (as figure
  \ref{fig:comp}A) for ChIP-Seq. The latest available version is in
  \url{https://github.com/welch16/ChIPUtils}.
\end{itemize}

\section{Conclusions}
\label{sec:conclusions}

We provide a ChIP-exo QC pipeline capable of assess the balance between
enriched samples and low complexity regions. It is shown that the
``peak-pair'' assumption doesn't hold in practice and we provide two
out-of-the-box visualization capable to assess the strand imbalance in
a ChIP-exo experiment.

We updated the dPeak algorithm from \cite{dpeak} and showed that
ChIP-exo is comparable in resolution and sensitivity with PET ChIP-Seq
and outperforms SET ChIP-Seq. ChIP-exo compared with PET and SET
ChIP-Seq at fixed depth sample is able to identify more targets at a
lower resolution. We compared dPeak with another algorithms to
estimate binding locations in ChIPexo data, dPeak is comparable to
MACE and outperforms GEM in resolution. dPeak provided a striking
balance in sensitivity, specificity and spatial resolution for
ChIP-exo analysis.

\newpage

\section{Planned work}
\label{sec:future}

\subsection{Some additional thoughts for this analysis}

\subsubsection{GC content and mappability quality indicators}
\label{sec:gc_map_qc}

Figure \ref{fig:comp} shows a relationship between the ChIP tag counts
of a ChIP-exo experiment and both mappability and GC content
scores. Furthermore, in sections \ref{sec:reco} and
\ref{sec:dpeak_analysis} we called peaks by using MOSAiCS which
considers them as necessary information when there is not input
sample. Hence, quality measures we may consider to add indicators to
the QC pipeline that considers this biases.

\subsubsection{Quality control for ChIP-Nexus}
\label{sec:nexus}

He et al., 2014 \cite{chipnexus} proposed a modification to the
ChIP-exo protocol that may fix its low library complexity issues by
extending the fragments by a random barcode prior to the exonuclease
digestion. This last statement is briefly mentioned by Mahony and
Pugh, 2015 \cite{exo_review}. We want to apply our QC pipeline to
ChIP-Nexus. Depending if ChIP-Nexus data shows background we may need
to modify the QC pipeline in the step where it paritions the reads
into regions, but the $\mbox{local-NSC}$ may be used to prove this
statement.

\subsection{Analysis of E. Coli Transcription Initiation Complexes}
\label{sec:ecoli}

The data showed in figure \ref{fig:exo_example} is actually part of a
more interesting experiment. We have 2 biological replicates of the
$\sig$, $\beta'_f$ and $\beta$ transcription factors under two
conditions, one where rifampicin was applied by 20 minutes and another
where it wasn't applied. As seen in the figure, there is ChIP-exo data
for the 3 factors and PET ChIP-Seq for $\sig$ and $\beta'_f$ (in the
figure both replicates were pooled).

$\sig$ factor is a transcription initiation factor of housekeeping
genes in E. Coli, and both $\beta$ factors transcribe the DNA into
mRNA. Hence, we want to use this data to have a better understanding
of the transcription in E. Coli.

The steps in formation of productive transcription complexes are:

\begin{enumerate}
\item Binding of RNAP to rpomoter DNA to form a closed complex.
\item Rearrangement of RNAP-DNA contacts to form an open complex.
\item Binding of nucleoside triphosphates in the active site and
  synthesis of a initial RNA to yield an initial transcription complex
  (ITC) - this is sometimes called a ``scrunched complex''.
\item Promoter escape to form a elongated complex (EC$_1$) when the
  promoter contacts are broken, and the $\sig$ factor is realeased.
\item Conversion of EC$_1$ into EC$_2$ by loading of the ribosome and
  elongation factors to enable efficient transcript elongation.
\end{enumerate}

% \begin{figure}[H]
% \centering
% \begin{tikzpicture}[node distance = 3cm]
% \tikzstyle{region} = [draw,thin,fill = blue!20]
% \tikzstyle{line} = [draw,thick]
% %\path[use as bounding box] (-1,0) rectangle (10,-2);
% \path[-] node[region] (step1) {RNAP + Promoters};
% \path[-] node[region,right of =step1,fill = red!20] (step2) {Closed};
% \path[-] node[region,right of =step2,fill = green!20] (step3) {Intermediate};
% \path[-] node[region,right of =step3,fill = yellow!20] (step4) {Open};
% \path[line][<->] (step1) -- (step2);
% \path[line][<->] (step2) -- (step3);
% \path[line][<->] (step3) -- (step4);
% \end{tikzpicture}
% \caption{Zooming into transcription}
% \end{figure}

As a proof of concept we were able to identify open and closed regions
by centering the binding the peaks respecto to the highest $\sig$
summit (among the rif0 and rif20 samples), and then using hierachical
clustering on the $\beta'_f$ signal of the rif0 samples to produce:

\begin{figure}[H]
  \centering
 \includegraphics[width = \textwidth]{/p/keles/ChIPexo/volume5/OpenClosedComplexes/figs/filtered/heatmaps_centroid.png}
 \caption{Centered and normalized heatmaps for PET ChIP-Seq peaks of
   $\sig$ and $\beta'_f$ under rif0 and rif20. The regions are
   organized by the centroids clustering of the $\beta'_f$-rif0
   profiles.}
  \label{fig:hm1}
\end{figure}

Figure \ref{fig:hm1} shows heatmaps of the normalized coverages of PET
ChIP-Seq peaks for $\beta'_f$ and $\sig70$ under rif0 and rif20
treatments. Each row represents the same region, under the four
different samples. As expected the signal for both $\sig$ samples is
centered around the middle, for $\beta'_f$-rif20, the majority of he
signal seems to be centered, hence the trascription was interrupted
and for $\beta'_f$-rif0 we can see clusters formed by open regions.

Then, using a set of high quality PET ChIP-Seq peaks and clustering
with the mclust model by Fraley and Raftery, 2007 \cite{mclust} we
were able to label the a set of regions as closed or open
complexes. Hence, we expect that by using ChIP-exo data we may find
different patterns that represent transcription units whose output was
limited at differemt steps in the formation of EC$_2$.

Additionally if we find several cases of those patterns, then we may
be able to determine which sequences may be causing this behaviour at
the different steps.

\subsection{Enhancer prediction and active learning}
\label{sec:enhancer}

Last summer, we entered to a contest from ENCODE's functional
characterization group to fit a model to classify enhancers. Roughly
speaks, the data set consided of a set of regions
$\mathcal{R}_1,\cdots,\mathcal{R}_n$ and a set of labels
$Y_1,\cdots,Y_n \in \{0,1,\}$. Hence, we built a series of features
$X_1,\cdots,X_p$ and fitted a model by dividing the $n$ observations
into a training set and a test set, etc. However, there are in total
another $m >> n$ regions in the experiment, and the cost of labelling
those $m$ regions may be high (in either the actual monetary cost of
the experiment or the time the scientists spends labelling those
regions). Those $n+m$ regions are H3K27ac peaks, hence they were
ranked and the first $n$ were selected randomly from groups with high,
medium and low signal. This raises the following questions:

\begin{itemize}
\item Is there a better way of ranking this $n+m$ regions ?
\item What is the best way to select the regions that are going to be
  labelled, i.e. we would want to label the regions that gives us the
  best model.
\end{itemize}

Hence, it may be worth to investigate which active learning strategies
may be the best to apply in this particular problem. Settles, 2012
\cite{al} is a survey of acive learning techniques that we want to
explore in order to solve this problems.

\newpage

\bibliographystyle{plain} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{chip_exo_paper}
\nocite{exo_gb}
\nocite{maplot1}
\nocite{maplot2}
\nocite{esl}
\nocite{anp}
\nocite{al}
\nocite{oc_review}
\nocite{exo4}

\section*{Supplement}
\label{sec:supp}

\subsection*{Additional Enrichment plots for $\sigma^{70}$}
\label{sec:enrichsup}

In figure \ref{fig:enrich2} it is shown the same relationship as in
figure \ref{fig:enrich}, but considering only regions formed where the
reads are being allocated in more than 10 (A) and 30 (B) unique
positions respectively. Both plots show that the vertical arms formed
by regions with low $\mbox{ARC}$ is formed by low complexity regions,
that way suggesting that this segment correspond to the background of
a ChIP-exo experiment.

\begin{figure}[h!]
  \centering
  \includegraphics[width = .8\textwidth,page = 3]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
  \includegraphics[width = .8\textwidth,page = 4]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
  \caption{Hexbin plots of $\mbox{ARC}$ vs $\mbox{URCR}$ for each
    region after partitioning the genome. In A and B the regions with
    reads mapped to at most 10 and 30 positions respectively where not
    considered. $\mbox{ARC}$ is defined as the ratio of the nr. of
    reads and the width of a region and $\mbox{URCR}$ is the ratio of
    the number of unique position where the reads are being allocated
    and the number of reads in a region.}
  \label{fig:enrich2}
\end{figure}

\end{document}