### 2015-12-01 rmarkdown/slides.Rnw
finished slides for prelim

\documentclass[compress,table,xcolor=dvipsnames]{beamer}

\usepackage{xcolor}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usepackage{float}
\usepackage{verbatim}
\usepackage{hyperref}

\usepackage{wrapfig}

\usepackage{multirow}

\usepackage{lmodern,tabularx,ragged2e,booktabs}
\newcolumntype{Y}{>{\arraybackslash\RaggedRight}X}
\newcolumntype{P}[1]{>{\arraybackslash\RaggedRight}p{#1}}

\usetheme{Szeged}
\usecolortheme[named=RoyalBlue]{structure}
\setbeamertemplate{headline}{}
\setbeamertemplate{navigation symbols}{}

% Personal Data
\def\Put(#1,#2)#3{\leavevmode\makebox(0,0){\put(#1,#2){#3}}}

\title[QC and analysis of ChIP-exo]{Data exploration, quality control
  and statistical analysis of ChIP-exo experiments}

\author{Rene Welch\\Preliminary Examination}

\institute[UW-STAT]{Department of Statistics\\University of Wisconsin - Madison}

\setbeamertemplate{section in toc}[sections numbered]
\setbeamertemplate{subsection in toc}[subsections numbered]

\setbeamertemplate{section in toc}[square]
\setbeamertemplate{subsection in toc}[square]

\date{December 1st, 2015}

\newcommand{\sig}{\sigma^{70}}

\begin{document}

\begin{frame}
  \maketitle
\end{frame}

\begin{frame}
\frametitle{Outline}
  \tableofcontents
\end{frame}

\section{Background}

\begin{frame}
\frametitle{ChIP-exo procedure}
  \begin{figure}[H]
    \centering
\includegraphics{../figs/diagrams/ChIPexo_diagram1.png}
\includegraphics{../figs/diagrams/ChIPexo_diagram2.png}\\
\includegraphics{../figs/diagrams/ChIPexo_diagram3.png}
    \caption{ChIP-exo procedure, Furey, 2012 \cite{chipbeyond}}
    \label{fig:exo_diagram}
  \end{figure}
\end{frame}

\section{ChIP-exo Data Exploration and Quality Control}

\subsection{ChIP-Seq QC measures}

\begin{frame}[t,plain]
  \frametitle{ChIP-Seq QC measures}

  \begin{itemize}
\only<1>{
\item Number of reads. Self-explanatory, the higher the better
  (usually).}
  \item<2-> Number of reads.
\only<2>{
\item PCR bottleneck Coeff. Ratio of number of pos. to which
  {\color{RoyalBlue}EXACTLY} one reads maps and number of pos. to
  which {\color{RoyalBlue}AT LEAST} one reads maps.}
\only<2>{
\begin{figure}[H]
  \centering
  \includegraphics[width = .75\textwidth]{../figs/for_paper/PBC_example.pdf}
\end{figure}
}
\item<3-> PCR bottleneck Coeff.
\only<3,4>{
\item Strand Cross-Correlation. 
\begin{align}
  y(\delta) = \sum_c w_c r\left[ n_c^+ \left(x + \frac{\delta}{2}
    \right), n_c^- \left( x- \frac{\delta}{2} \right)\right]
\nonumber
\end{align}
}
\only<3>{
  \begin{figure}[H]
    \centering
    \includegraphics[width = .4\textwidth]{../figs/for_paper/Kharchenko.png}
\caption{SCC explanation. Kharchenko et al., 2008 \cite{strandcc}}
  \end{figure}
}
\only<4>{
\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth]{../figs/for_paper/SCC.png}
\caption{SCC as QC. Landt et al., 2012 \cite{encode_qc}}
\end{figure}
}
\item<5-> Strand Cross-Correlation.
\only<5>{
\item Normalized Strand Cross-Correlation. Ratio between the SCC when
  the shift is the fragment length and min. value of the SCC.
}
\item<6> Normalized Strand Cross-Correlation. 
\only<6>{
\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth]{../figs/for_paper/SCC.png}
\caption{SCC as QC. Landt et al., 2012 \cite{encode_qc}}
\end{figure}
}
\end{itemize}
\end{frame}

\subsection{Evaluation of ChIP-Seq QC Metrics for ChIP-exo}
\begin{frame}[plain]
\frametitle{Evaluation of ChIP-Seq QC Metrics for ChIP-exo}

\setlength\tabcolsep{3pt}  % default value: 6pt
\scriptsize  %%  command to change the font size
% \rowcolor{RoyalBlue!20} %-- this indicates the change in odd and pair rows
\begin{table}[H]
\begin{tabularx}{\textwidth}{ @{} c|c|c|c|c|c|c| Y @{}}
  \toprule
\textbf{IP} & \textbf{Organism}  & \textbf{Condition} & \textbf{Rep.} & \textbf{Nr. reads} & \textbf{PBC} &  \textbf{NSC} & \textbf{Source} \\
\midrule
$\sig$ & E.Coli & Rif-0min & 1 & 960,256 & 0.2823 &   10.29   & \multirow{4}{\hsize}{Courtesy of Prof. Landick's lab} \\
$\sig$ &  E.Coli & Rif-0min & 2 & 2,247,295 & 0.2656 &  25.08  \\
$\sig$ &  E.Coli & Rif-20min & 1 & 1,940,387 & 0.2698 &  17.69  \\
$\sig$ &  E.Coli & Rif-20min & 2 & 4,229,574 & 0.2153 &   14.11 \\
\hline
FoxA1 &  Mouse &  - & 1 & 22,210,461 & 0.6562 & 21.452 & \multirow{3}{\hsize}{From Serandour et al., 2013 \cite{exoillumina}} \\
FoxA1 &  Mouse &  - & 2 & 22,307,557 & 0.7996 & 60.661 \\
FoxA1 &  Mouse &  - & 3 & 22,421,729 & 0.1068 & 72.312 \\
\hline
ER & Human & - & 1 & 9,289,835 & 0.8082 & 19.843 & \multirow{3}{\hsize}{From Serandour et al., 2013 \cite{exoillumina}} \\
ER & Human & - & 2 & 11,041,833 & 0.8024 & 21.422 \\
R & Human & - & 3 & 12,464,836 & 0.8203 & 19.699 \\
\hline
CTCF & Human & - & 1 &   48,478,450 & 0.4579 & 15.977 & From Rhee and Pugh 2011, \cite{exo1} \\
\bottomrule
\end{tabularx}  
\end{table}

{\small
\begin{itemize}
\item For PBC (human and mouse): 0 - 0.5 severe bottlenecking , 0.5 -
  0.8 moderate bottlenecking, 0.8 - 0.9 mild bottlenecking, 0.9 - 1 no
  bottlenecking.
\item For NSC (human and mouse): $ < 1.1$ is relatively low.
\end{itemize}
}
\end{frame}

\subsection{Comparison of ChIP-exo and ChIP-seq}

\begin{frame}
  \frametitle{Comparison of ChIP-exo and ChIP-Seq}

\begin{figure}[H]
  \centering
  \includegraphics[width = .5\textwidth,page = 3 ]{../figs/for_paper/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
  \includegraphics[width = .5\textwidth]{../figs/for_paper/forward_strand_ratio_comp_old.pdf}
\end{figure}

{\small
\begin{itemize}
\item A) shows that high density regions are similar between ChIP-Seq
  and ChIP-exo but background regions are not.
\item The peak-pair assumption does hold in ChIP-exo data, but not
  locally since some regions show strand-imbalance.
\end{itemize}
}

\end{frame}

\begin{frame}[t]
  \frametitle{Comparison of ChIP-exo and ChIP-seq}  
\begin{columns}
\column{.7\textwidth}
\begin{figure}[H]
  \centering  
\includegraphics[width = .9\textwidth]{../figs/for_paper/scc_ctcf.pdf}
\end{figure}

\column{.3\textwidth}
  \begin{minipage}[c][.6\textheight][c]{\linewidth}
{\small
  {\color{RoyalBlue}Figure:} SCC for CTCF factor in HeLa cell line for
    ChIP-exo and SET-ChIP-Seq.\\
    - {\color{ForestGreen} PBC for rep1 is 0.56}
    - {\color{ForestGreen} PBC for rep2 is 0.94}
  }
\end{minipage}
\end{columns}

{\small
\begin{itemize}
\item There is a \emph{``phantom peak''} at read length for ChIP-Seq (SE).
\item In ChIP-Seq, SCC is maximized at the unobserved fragment length.
\item In ChIP-exo, the \emph{``phantom peak''} and the fragment length
  summit are confounded.
\end{itemize}
}

\end{frame}

\subsection{Quality Control Pipeline}

\begin{frame}[t]
  \frametitle{Quality Control Pipeline}

  \begin{enumerate}
\only<1>{
  \item Partition the genome and generate ChIP-exo islands.
\begin{figure}[H]
\centering
\includegraphics[width = \textwidth]{../figs/for_paper/coverage_diagram2_part1.pdf}
\end{figure}
}
\only<2>{
\addtocounter{enumi}{1}
\item Calculate a vector of summary statistics for each islands.
\begin{figure}[H]
\centering
\includegraphics[width = \textwidth]{../figs/for_paper/coverage_diagram2_part2.pdf}
\end{figure}
}
\only<3>{
\addtocounter{enumi}{2}
\item Visualize all islands together.
\begin{figure}[H]
\centering
\includegraphics[width = \textwidth]{../figs/for_paper/coverage_diagram2_part3P.pdf}
\end{figure}
}
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{ARC vs. URCR exploration}

\only<1-2>{
  \begin{figure}[H]
    \centering
    \includegraphics[width = .7\textwidth,page =1]{../figs/for_paper/enrichment_example.pdf}
  \end{figure}
}

\only<2>{
\Put(120,200){\includegraphics[width = .6\textwidth,page = 1]{../figs/for_paper/coverage_example.pdf}}
}

\only<3-4>{
  \begin{figure}[H]
    \centering
    \includegraphics[width = .7\textwidth,page =2]{../figs/for_paper/enrichment_example.pdf}
  \end{figure}
}

\only<4>{
\Put(120,200){\includegraphics[width = .6\textwidth,page = 2]{../figs/for_paper/coverage_example.pdf}}
}

\only<5-6>{
  \begin{figure}[H]
    \centering
    \includegraphics[width = .7\textwidth,page =3]{../figs/for_paper/enrichment_example.pdf}
  \end{figure}
}

\only<6>{
\Put(120,200){\includegraphics[width = .6\textwidth,page = 3]{../figs/for_paper/coverage_example.pdf}}
}

\only<7-8>{
  \begin{figure}[H]
    \centering
    \includegraphics[width = .7\textwidth,page = 4]{../figs/for_paper/enrichment_example.pdf}
  \end{figure}
}

\only<8>{
\Put(120,200){\includegraphics[width = .6\textwidth,page = 4]{../figs/for_paper/coverage_example.pdf}}
}

\only<9-10>{
  \begin{figure}[H]
    \centering
    \includegraphics[width = .7\textwidth,page =6]{../figs/for_paper/enrichment_example.pdf}
  \end{figure}
}

\only<10>{
\Put(120,300){\includegraphics[width = .6\textwidth,page = 6]{../figs/for_paper/coverage_example.pdf}}
}

\only<11-12>{
  \begin{figure}[H]
    \centering
    \includegraphics[width = .7\textwidth,page =7]{../figs/for_paper/enrichment_example.pdf}
  \end{figure}
}

\only<12>{
\Put(120,300){\includegraphics[width = .6\textwidth,page = 7]{../figs/for_paper/coverage_example.pdf}}
}

\only<13-14>{
  \begin{figure}[H]
    \centering
    \includegraphics[width = .7\textwidth,page =8]{../figs/for_paper/enrichment_example.pdf}
  \end{figure}
}

\only<14>{
\Put(120,300){\includegraphics[width = .6\textwidth,page = 8]{../figs/for_paper/coverage_example.pdf}}
}

\end{frame}

\begin{frame}
  \frametitle{ARC vs. URCR exploration by replicate}
\only<1>{
All regions:
\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth,page = 1]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
\end{figure}
}

\only<2>{
Regions formed by reads aligned to more than 10 unique positions:
\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth,page = 3]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
\end{figure}
}

\only<3>{
Regions formed by reads aligned to more than 30 unique positions:
\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth,page = 4]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
\end{figure}
}

\end{frame}

\begin{frame}
  \frametitle{ARC vs. URCR and local-NSC}

\only<1>{
Calculate coverage for all regions:
\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth,page = 1]{../figs/for_paper/local_SCC_separated.pdf}
\end{figure}
}

\only<2>{
Calculate $\mbox{local-SCC}$ for regions:
\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth,page = 2]{../figs/for_paper/local_SCC_separated.pdf}
\end{figure}
}

\only<3>{
Fit local polynomial regression for the $\mbox{local-SCC}$ :
\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth,page = 3]{../figs/for_paper/local_SCC_separated.pdf}
\end{figure}
}

\only<4>{
\begin{align}
 y(\delta) = f(x_\delta) + \epsilon_\delta, \quad 
\mbox{local-NSC} = \frac{max_{x_\delta} \hat{f}(x_\delta)}{\hat{\sigma}_f},
\nonumber
\end{align}
\vspace*{\fill}
}

\only<4>{
\begin{columns}

\column{.4\textwidth}
where:

\begin{itemize}
\item $x_\delta$ is the shift.
\item $y_\delta$ is the local - SCC at shift $x_\delta$.
\item $\epsilon \sim N(0, \sigma^2)$ i.i.d .
\end{itemize}

\column{.6\textwidth}
{\scriptsize
  \begin{table}[H]
    \centering
    \begin{tabularx}{\textwidth}{ @{} l | r | r |r @{}}
      \toprule
      Replicate & $\hat{\sigma}$ & $\max_{x_\delta} \hat{f}(x_\delta)$ & $\mbox{local-NSC}$ \\
      \midrule
Rep. 1 & 0.0533 & 0.2215 & 4.14 \\
Rep. 2 & 0.0453 & 0.1606 & 3.54 \\
Rep. 3 & 0.0475 & 0.0832 & 1.75 \\
\bottomrule
    \end{tabularx}
  \end{table}
}

\end{columns}
}

\only<5>{
  \begin{figure}[H]
    \centering
      \includegraphics[width = \textwidth,page =4]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata.pdf}
  \end{figure}

  \begin{itemize}
  \item {\color{RoyalBlue} high:} Number of unique positions $ > 100$.
  \item {\color{RoyalBlue} med:} Number of unique positions in $(50,100)$.
  \item {\color{RoyalBlue} low:} Number of unique positions in $(20,50)$.
  \end{itemize}
}
\end{frame}

\begin{frame}[t]
  \frametitle{Strand imbalance}

\begin{figure}[H]
  \centering  
  \includegraphics[width = \textwidth,page = 3]{../figs/Carroll_mice_for_paper/Strand_imbalance.pdf} 
  \label{fig:strand}
\end{figure}

{\scriptsize
\begin{itemize}
\only<1>{
\item The quantiles of the first replicate, approximate the median
  faster than the other replicate's quantiles.
 }
\only<2>{
\item B) shows that the majority of the low depth islands are formed
  exclusively by one strand.
}

\end{itemize}
}

\end{frame}

\section{Systematic Evaluation of ChIP-Seq and ChIP-exo}

\begin{frame}[plain]
  \frametitle{dPeak model for SET case}

\only<1>{

      \begin{figure}[H]
        \centering
        \includegraphics[width =
        \textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/poster/dpeak.png}
      \end{figure}
}

\only<2>{
\begin{columns}[t]
\column{.3\textwidth}
  \begin{figure}[H]
    \centering
    \includegraphics[width =\textwidth]
{/p/keles/ChIPexo/volume3/ChIPexo/poster/dpeak.png}
  \end{figure}

\column{.7\textwidth}
  We consider a region with $n$ reads and $m$ positions, for the
  $i$-th read:
  \begin{itemize}
  \item $Z_i \sim \mbox{Multi}(\pi_0,\pi_1,\cdots,\pi_{g^*})$
  \item $D_i \sim \mbox{Ber}(p_D)$
    \begin{itemize}
    \item The read is in the forward strand ($D_i = 1$):
      \begin{itemize}
      \item The reads belongs to the background: $R_i | Z_i = 0, D_i =
        1 \sim \mbox{Unif}(1, m)$
      \item The read belong to the $g$-th binding event: $R_i | Z_i =
        g, D_i = 1 \sim \mbox{N}(\mu_g - \delta , \sigma^2)$
      \end{itemize}
    \item The read is in the backward strand ($D_i = 0$):
      \begin{itemize}
      \item The reads belongs to the background: $R_i | Z_i = 0, D_i =
        0 \sim \mbox{Unif}(1 , m )$
      \item The read belong to the $g$-th binding event: $R_i | Z_i =
        g, D_i = 0 \sim \mbox{N}(\mu_g + \delta , \sigma^2)$
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{columns}
}

\end{frame}

\begin{frame}[plain]
  \frametitle{Data structure}
Available for $\sig$ (courtesy of Prof. Landick's lab):
  \begin{figure}[H]
    \centering
    \includegraphics[width = \textwidth]{../figs/for_paper/chip_explanation.pdf}
  \end{figure}

\end{frame}

\begin{frame}
\frametitle{Comparison with ChIP-Seq using dPeak}
\only<1>{
  \begin{figure}[H]
  \centering
  \includegraphics[width = .46\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
  \includegraphics[width = .46\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
\end{figure}
}
\only<1>{
{\small
  \begin{itemize}
  \item Sensitivity is the defined as the proportion of identified
    peaks (regulonDB \cite{regulonDB} is used as gold-standard)
  \item Resolution is defined as the min. absolute distance of a
    regulonDB annotation to an est. binding location.
  \end{itemize}
}}

\only<2>{
\begin{figure}[H]
\centering
   \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/sigma_delta_old_densities.pdf}
   \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/sigma_delta_old_densities.pdf} 
 \end{figure}
}
\only<2>{
{\small
\begin{itemize}
\item $\delta$ measures the average distance of reads to their
  respective binding sites.
\item $\sigma$ measures the dispersion of reads around their
  respective binding sites.
\item Good news !! the model is reflecting the truth.
\end{itemize}
}}
\end{frame}

\begin{frame}[t]
  \frametitle{ChIP-Seq comparison at fixed depth}

  We sampled $n$ fragment reads of each dataset ($2n$ for PET
  ChIP-Seq), and applied the MOSAiCS / dPeak pipeline:

\only<1>{
\begin{figure}[H]
  \centering
  \includegraphics[width = .4\textwidth,page = 1]{../figs/for_paper/saturation_analysis_old_frames.pdf}
  \includegraphics[width = .4\textwidth,page = 2]{../figs/for_paper/saturation_analysis_old_frames.pdf}
\end{figure}
}

\only<1>{
{\small
  \begin{itemize}
  \item There are more candidate regions for both ChIP-Seq samples,
    hence there are also more predicted events.
  \end{itemize}
}}

\only<2>{
\begin{figure}[H]
\centering
\includegraphics[width = .4\textwidth,page = 3]{../figs/for_paper/saturation_analysis_old_frames.pdf}
  \includegraphics[width = .4\textwidth,page = 4]{../figs/for_paper/saturation_analysis_old_frames.pdf}
\end{figure}
}

\only<2>{
{\small
  \begin{itemize}
  \item ChIP-exo and PET ChIP-Seq are comparable and outperform SET
    ChIP-Seq
  \end{itemize}
}}

\end{frame}

\section{Comparison with Other Algorithms}
\begin{frame}
\frametitle{Comparison with Other Algorithms}

\begin{figure}[H]
  \centering
  \includegraphics[width = .6\textwidth]{../figs/for_paper/algorithm_resolution.pdf}
\end{figure}

{\scriptsize
  \begin{itemize}
  \item Resolution is defined as the min. absolute distance from the
    regulonDB annotation to an estimated binding site.
  \end{itemize}
}

\end{frame}

\section{Conclusions and future work}

\begin{frame}
  \frametitle{Conclusions}

  \begin{itemize}
  \item Our pipeline is capable of assessing the balance between
    sample enrichment and library complexity.
  \item We shown that the ``peak-pair'' assumption does not hold well
    in practice, and implemented a visualization capable of detecting
    strand imbalance.
  \item We updated dPeak, which makes a striking balance in
    sensitivity, specificity and spatial resolution.
  \item ChIP-exo and PET ChIP-Seq are comparable in resolution and
    sensitivity, and both outperform SET ChIP-Seq.
  \item We showed that with a fixed number of reads, ChIP-exo
    outperforms PET and SET ChIP-Seq.
  \item dPeak outperforms other algorithms in resolution.
  \end{itemize}
\end{frame}

\begin{frame}[label=future]
\frametitle{Future work}

\begin{itemize}
\item In the paper, we showed that there is a relationship between
  ChIP-exo tag counts and both mappability and GC content scores. We
  want to add a QC measure to the pipeline based on
  them. \hyperlink{gcmap}{\beamerbutton{GC \& Mapp}}
\item We want to assess if ChIP-Nexus library complexity is actually
  higher than ChIP-exo's by using the
  $\mbox{local-NSC}$. \hyperlink{nexus}{\beamerbutton{ChIP-Nexus}}

\item We have been studying E. Coli's transcription initiation
  complexes with PET ChIP-Seq. We want to improve this analysis by
  using ChIP-exo data.  \hyperlink{tic}{\beamerbutton{TIC}}
\item Find a optimal strategy for labeling enhancer out of a
  predetermined list of regions in the genome by the use of active
  learning techniques.
\end{itemize}

\end{frame}

\begin{frame}[t]
\frametitle{Software}  
\begin{itemize}
\item {\color{RoyalBlue}\textbf{dPeak}}: We updated the initialization
  strategy. The latest version is currently available from
  \url{http://dongjunchung.github.io/dpeak/}.
\item {\color{RoyalBlue}\textbf{ChIPexoQual}}: This package contains
  the QC pipeline for ChIP-exo. The last version is available in
  \url{https://github.com/welch16/ChIPexoQual}.
\item {\color{RoyalBlue}\textbf{Segvis}}: The goal of this package is
  to visualize genomic regions by using aligned reads. The latest
  version is available in \url{https://github.com/keleslab/Segvis}.
\item {\color{RoyalBlue}\textbf{ChIPUtils}}: This package attempts to
  gather the most commonly used ChIP-Seq QC.The latest available
  version is in \url{https://github.com/welch16/ChIPUtils}.
\end{itemize}

\end{frame}

\begin{frame}[plain]

{\Huge
{\color{RoyalBlue}
\begin{center}
  \textbf{Thank you very much!}
\end{center}
}
}

\end{frame}

\begin{frame}[allowframebreaks]
  \frametitle{References}

{\tiny

\nocite{exo1}
\nocite{dpeak}

\bibliographystyle{plain} % Style BST file (bmc-mathphys, vancouver, spbasic)
\bibliography{chip_exo_paper}

}

\end{frame}

\begin{frame}[label=gcmap]
\frametitle{GC content and Mappability score bias}

\begin{figure}[H]
  \centering
  \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
  \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}

  \begin{itemize}
  \item ChIP tag counts increase linearly as mappability scores
    increases.
  \item ChIP tag counts increase linearly as GC content score
    increases when GC content is less than 0.6 and then ChIP tag
    counts decrease as GC content
    increases. \hyperlink{future}{\beamerbutton{back}}

  \end{itemize}

\end{figure}

\end{frame}

\begin{frame}[plain,label=nexus]
\frametitle{ChIP-exo and ChIP- Nexus}

\begin{columns}
\column{.5\textwidth}
  \begin{figure}[H]
    \includegraphics[width = .9\textwidth]{../vignettes/exo-protocol.png}
\caption{ChIP-exo explanation, Mendenhall and Bernstein, 2012 \cite{exo_gb}}
  \end{figure}
\column{.5\textwidth}

\begin{itemize}
\item ChIP-Nexus uses a random barcode for adapter $\#$ 1.
\item Before step 5 uses self-circularization to attach the random
  barcode to the 3' end, and uses it instead of adapter $\#$ 2.

\end{itemize}
  \begin{figure}[H]
    \centering
    \includegraphics[width = \textwidth]{../figs/for_paper/nexus.png}
    \caption{ChIP-Nexus example, He et al., 2015, \cite{chipnexus}
      \hyperlink{future}{\beamerbutton{back}}}
  \end{figure}

\end{columns}

\end{frame}

\begin{frame}[label=tic]
\frametitle{Transcription Initiation Complexes}

  \begin{figure}[H]
  \centering
 \includegraphics[width = \textwidth]{/p/keles/ChIPexo/volume5/OpenClosedComplexes/figs/filtered/heatmaps_centroid.png}
 \caption{Centered and normalized heatmaps for PE ChIP-Seq peaks of
   $\sig$ and $\beta'_f$ under rif0 and rif20. The regions are
   organized by the centroids clustering of the $\beta'_f$-rif0
   profiles. \hyperlink{future}{\beamerbutton{back}}}
  \label{fig:hm1}
\end{figure}

\end{frame}

\end{document}

% LocalWords:  regulonDoB