### 2016-03-21 rmarkdown/paper.tex
updated paper figures and fixed format

% \VignetteEngine{knitr::knitr}
\documentclass{bmcart}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

\usepackage{xcolor}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usepackage{float}
\usepackage{verbatim}
% \usepackage{hyperref}
% \hypersetup{
%      colorlinks   = true,
%      citecolor    = black
% }
% \hypersetup{linkcolor=black}
% \hypersetup{linkcolor=black}
% \hypersetup{linkcolor=black}

\usepackage{wrapfig}

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

% \usepackage{authblk}
\usepackage[utf8]{inputenc} %unicode support

\newcommand{\sig}{\sigma^{70}}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

%\maketitle

\begin{frontmatter}

\begin{fmbox}
\dochead{Draft}

\title{Data exploration, quality control and statistical analysis of
  ChIP-exo experiments}

\author[
   addressref={aff1},
   noteref={n1},
   email={welch@stat.wisc.edu}   
]{\inits{RW}\fnm{Rene} \snm{Welch} }
\author[
   addressref={aff6},        % id's of addresses, e.g. {aff1,aff2}
   noteref={n1},
   email={chungd@musc.edu}   % email address
]{\inits{DC}\fnm{Dongjun} \snm{Chung} } 
\author[
   addressref={aff3},
   email={ong@cs.wisc.edu}
]{\inits{IO}\fnm{Irene} \snm{Ong}}
\author[
   addressref={aff3,aff4},
   email={jagrass@wisc.edu}
]{\inits{JG}\fnm{Jeffrey} \snm{Grass}}
\author[
   addressref={aff3,aff4,aff5},
   email={landick@bact.wisc.edu}
]{\inits{RL}\fnm{Robert} \snm{Landick}}
\author[
   addressref={aff1,aff2},
   corref={aff1},
   email={keles@stat.wisc.edu}
]{\inits{SK}\fnm{S\"und\"uz} \snm{Kele\c{s}}}

\address[id=aff1]{%                           % unique id
  \orgname{Department of Statistics, University of Wisconsin Madison}, % university, etc
  \street{1300 University Avenue},                     %
  %\postcode{}                                % post or zip code
  \city{Madison},                              % city
  \cny{WI}                                    % country
}

\address[id=aff2]{%
  \orgname{Department of Biostatistics and Medical Informatics, University of Wisconsin Madison},
  \street{600 Highland Avenue},
%  \postcode{24105}
  \city{Madison},
  \cny{WI}
}
\address[id=aff3]{
  \orgname{Great Lakes Bioenergy Research Center, University of Wisconsin Madison},
  \street{1552 University Avenue},
  \city{Madison},
  \cny{WI}
}
\address[id=aff4]{
  \orgname{Department of Biochemistry, University of Wisconsin Madison},
  \street{433 Babcock Drive},
  \city{Madison},
  \cny{WI}
}
\address[id=aff5]{
  \orgname{Department of Bacteriology, University of Wisconsin Madison},
  \street{1550 Linden Drive},
  \city{Madison},
  \cny{WI}
}
\address[id=aff6]{
  \orgname{Department of Public Health Sciences, Medical University of South Carolina},
  \street{135 Cannon Street},
  \city{Charleston},
  \cny{SC}
}

\begin{artnotes}
\note[id=n1]{These two authors contributed equally.} % note, connected to author
\end{artnotes}

\end{fmbox}

\begin{abstractbox}

  \begin{abstract}

    ChIP-exo is a modification of the ChIP-Seq protocol for high
    resolution mapping of transcription factor binding sites. Although
    many aspect of the ChIP-exo data analysis are similar to those of
    ChIP-Seq, ChIP-exo presents a number of unique challenges. We
    present a quality control pipeline that analyzes a ChIP-exo
    experiment's strand imbalance, enrichment and library
    complexity. Assessment of these characteristics are facilitated
    through diagnostic plots and summary statistics calculated over
    regions of the genome with varying levels of coverage.

    We systematically evaluated diverse aspects of ChIP-exo and found
    the following characteristics: First, ChIP-exo's background is
    quite different from ChIP-Seq's. Second, although often assumed in
    ChIP-exo data analysis methods, the ``peak pair'' assumptions does
    not hold locally in actual ChIP-exo data. Third, we for the first
    time compared Paired End (PE) ChIP-Seq with ChIP-exo and found
    that both protocols are comparable in resolutions and sensitivity
    for closely located binding events, but as the distance between
    binding events increases ChIP-exo shows higher sensitivity that PE
    ChIP-Seq. Finally, at fixed sequencing depths, ChIP-exo provides
    higher sensitivity, specificity and spatial resolution than PE
    ChIP-Seq.

  \end{abstract}

\begin{keyword}
  \kwd{ChIP-exo}
  \kwd{Quality Control}
  \kwd{ChIP-Seq}
  \kwd{Spatial Resolution}
  \kwd{Transcription Factor}  
  \kwd{Binding Site Identification on High-Res}
  \kwd{Deconvolution}
\end{keyword}

\end{abstractbox}

\end{frontmatter}

\newpage

\section{Background}
\label{sec:intro}

ChIP-exo (Chromatin Immunoprecipitation followed by exonuclease
digestion and next generation sequencing) Rhee and Pugh, 2011
\cite{exo1} is the state-of-the-art experiment developed to attain
single base-pair resolution of protein binding site identification and
it is considered as a powerful alternative to popularly used ChIP-Seq
(Chromatin Immunoprecipitation coupled with next generation
sequencing) assay. ChIP-exo experiments first capture millions of DNA
fragments (150 - 250 bp in length) that the protein under study
interacts with using random fragmentation of DNA and a
protein-specific antibody. Then, exonuclease is introduced to trim the
$5\prime$ end of each DNA fragment to a fixed distance from the bound
protein compared to ChIP-Seq. This step is unique to ChIP-exo and
could potentially provide significantly higher spatial resolution
compared to ChIP-Seq. Finally, high throughput sequencing of a small
region (36 to 100 bp) at the $5\prime$ end of each fragment generates
millions of reads. Figure \ref{fig:chip_diagram} illustrates the
differences between ChIP-exo, Single End (SE) ChIP-Seq and Paired End
(PE) ChIP-Seq: The $5\prime$ ends of a ChIP-exo experiment are located
more tightly around the binding proteins than in a ChIP-Seq
experiment; in a PE ChIP-Seq experiment both ends are observed while
in a SE ChIP-Seq experiment only the $5\prime$ end.

% Reffers to... First, DNA libraries... I think that these might be
% too much details for the introduction. How about discussing them in
% the results when we investigate the related

While the number of ChIP-exo data keeps increasing, characteristics of
ChIP-exo data are not fully investigated yet. First, DNA libraries
generated by the ChIP-exo protocol seem to be less complex than the
libraries generated by ChIP-Seq (Mahony et al., 2015
\cite{exo_review}). Second, although there are roughly the same amount
of reads in both strands, locally there may be more reads in one
strand than in the other. Finally, most of current ChIP-Seq quality
control (QC) guidelines (Landt et al., \cite{encode_qc}) may not be
applicable on ChIP-exo, while there are not established QC pipelines
for ChIP-exo; previous ChIP-exo analyses used ChIP-Seq samples to
compare the resolution between experiments (\cite{exo1},
\cite{exoillumina}, \cite{exo2})). To address these challenges, we
suggest a collection of quality control visualizations to interrogate
these biases in a ChIP-exo experiment and globally asses the
enrichment and library complexity of a ChIP-exo sample. We gathered
ChIP-exo data from diverse organisms: CTCF factor in human
\cite{exo1}; ER factor in human and FoxA1 factor in mouse (Serandour
et al., 2013 \cite{exoillumina}); Glucocorticoid receptor (GR) in
IMR90, K562 and U20S cell lines (Starick et al., 2015
\cite{starick15}).Furthermore, we also generated ChIP-exo and ChIP-Seq
data for $\sig$ factor in \emph{Escherichia Coli} (\emph{E. Coli})
measured under aerobic ($ + O_2$) condition, and treated by rifampicin
by 0 and 20 minutes (courtesy of Professor Robert Landick's lab).

In order to obtain the potential benefits of ChIP-exo on protein
binding site identification, it is critical to use algorithms that
could fully utilize information available in ChIP-exo data. Rhee and
Pugh, 2011 \cite{exo1} discussed that reads in the forward and reverse
strand might construct peak pairs around bound proteins, of which
heights were implicitly assumed to be symmetric. Based on this
rationale, they used the ``peak pair method'' that predicts the
midpoint of two modes of peak pairs as potential binding
sites. Recently developed ChIP-exo data analysis methods, such as Mace
(Wang et al, 20114 \cite{mace}), CexoR (Madrigal, 2015 \cite{cexor})
and Peakzilla (Bardet et al., 2013 \cite{peakzilla}), are also based
on this peak pair assumption. However, appropriateness of such
assumption was not fully evaluated in the literature yet. Furthermore,
it is still unknown which factors could affect protein binding site
identification using ChIP-exo data. In order to address this problem,
we investigated various aspects of ChIP-exo data by contrasting them
with their respective ChIP-Seq experiments.

Currently, research on statistical methods for ChIP-exo data is still
in its very early stage. Although many methods have been proposed to
identify protein binding sites from ChIP-Seq data (reviewed by
Wilbanks and Facciotti, 2012 \cite{evaluation} and Pepke and Wold,
2009 \cite{computation}), such as MACS (Zhang et al., 2008
\cite{macs}), CisGenome (Ji et al., 2008 \cite{cisgenome}) and MOSAiCS
(Kuan et al., 2009 \cite{mosaics}), these approaches might not fully
utilize potentials of ChIP-exo data for high resolution identification
of protein binding sites. Specifically these approaches reveal protein
binding sites only in lower resolution, i.e., at an interval of
hundreds to thousands of base pairs. Furthermore, they implicitly
assume that there is only one ``mode'' or ``predicted binding
location'' per this wide genomic interval. More recently,
deconvolution algorithms such as Deconvolution (Lun et al., 2009
\cite{csdeconv}), GEM (Guo et al., 2012 \cite{gem}, an improved
version of Guo et al., 2010 \cite{gps} ) and PICS (Zhang et al., 2010
\cite{pics}) have been proposed to identify binding sites in higher
resolution using ChIP-Seq data. However, most of them are still not
tailored for ChIP-exo and PE and SE ChIP-Seq data in a unified
framework and as a result, currently available methods are not
appropriate for fair comparison between ChIP-exo and ChIP-Seq. To
address these limitations, we developed and utilized an improved
version of dPeak (Chung et al., 2013 \cite{dpeak}), a high resolution
binding site identification (deconvolution) algorithm that we
previously developed for PE and SE ChIP-Seq data, so that it can also
handle ChIP-exo data. The dPeak algorithm implements a probabilistic
model that accurately describes the ChIP-exo and ChIP-Seq data
generation process.

Some of the key findings in this work are as follows. First, we
demonstrate that the ``peak pair'' assumption of Rhee and Pugh, 2013
\cite{exo2} does not hold well in real ChIP-exo data. Second, we found
that when we analyze ChIP-exo data from eukaryotic genomes, it is
important to consider sequence biases inherent to ChIP-exo data, such
as mappability and GC content, in order to improve sensitivity and
specificity of binding site identification. Third, we evaluated
several methods to identify binding events and dPeak performs
competitively respect to GEM and MACE when analyzing ChIP-exo
data. Finally, when comparable number of reads is used for both
ChIP-exo and ChIP-Seq , dPeak coupled with ChIP-exo data provides
resolution comparable to PE ChIP-Seq and both significantly improve
the resolution of protein binding identification compared to SE-based
analysis with any of the available methods.

\section{Results and discussion}
\label{sec:results}

\subsection{Deeply sequenced E. Coli $\sig$ ChIP-exo and ChIP-Seq data}

$\sig$ factor is a transcription initiation factor of housekeeping
genes in E. Coli. In this organism's genomes, many promoters contain
multiple transcription start sites (TSS) and these TSS are often
closely spaced (10 $\sim$ 150 bp). These closely spaced binding sites
are considered to be multiple ``switches'' that differentially
regulate gene expression under diverse growth conditions
\cite{regulondb}. Therefore, investigation of ChIP-exo's potential for
identification and differentiation of closely spaced binding sites is
invaluable for elucidating the transcriptional networks of prokaryotic
genomes.

\subsection{Application of current ChIP-Seq QC guidelines on ChIP-exo
  data}

We started our exploration by investigating whether the current
state-of-the-art QC pipelines for ChIP-Seq are suitable for
ChIP-exo. Table \ref{tab:qc} contains measures that are commonly
calculated for ChIP-Seq samples: PCR Bottleneck Coefficient (PBC) and
the Normalized Strand Cross-Correlation (NSC) are calculated as in
\cite{encode_qc}. We omitted the Relative Strand Cross-Correlation
(RSC) which is another commonly used QC measure because a typical
ChIP-exo experiment is not accompanied by an input sample.

%% We can observe that the PBC is low, hence it would be incorrectly
%% suggested to repeat the experiment.

\begin{table}[h!]
%% need to fix format
  \centering

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{verbatim}
##     Organism  IP/TF Condition/Cell Rep.       Depth    PBC      NSC
##  1:   E.Coli  Sig70        Aerobic    1  13,961,493 0.1399 103.4239
##  2:   E.Coli  Sig70        Aerobic    2  14,810,838 0.1634 162.8002
##  3:   E.Coli  Sig70      Anaerobic    1  16,108,774 0.1354 153.5088
##  4:   E.Coli  Sig70      Anaerobic    2  13,636,541 0.1532 172.5815
##  5:   E.Coli  Sig70       Rif-0min    1     902,921 0.2690  13.7691
##  6:   E.Coli  Sig70       Rif-0min    1   1,852,124 0.2591  17.9188
##  7:   E.Coli  Sig70      Rif-20min    2   2,104,427 0.2584  29.6083
##  8:   E.Coli  Sig70      Rif-20min    2  11,548,572 0.1511  13.0863
##  9:    Human   CTCF           HeLA    1  48,478,450 0.4580  16.0248
## 10:    Mouse Fox A1    Mouse Liver    1  22,210,461 0.6562  21.2820
## 11:    Mouse Fox A1    Mouse Liver    2  23,307,557 0.7996  60.4219
## 12:    Mouse Fox A1    Mouse Liver    3  22,421,729 0.1068  72.0424
## 13:    Human     GR          IMR90    1  47,443,803 0.2979   8.8602
## 14:    Human     GR           K562    1 116,518,000 0.0505   4.1179
## 15:    Human     GR           U20S    1   3,255,111 0.7714  10.0588
## 16:    Human     ER          MCF-7    1   9,289,835 0.8083  19.8752
## 17:    Human     ER          MCF-7    2  11,041,833 0.8024  21.4851
## 18:    Human     ER          MCF-7    3  12,464,836 0.8204  18.7281
\end{verbatim}
\end{kframe}
\end{knitrout}
\caption{\textbf{Current QC metrics applied to ChIP-exo data.} PBC
  stands for PCR Bottleneck Coefficient and NSC for Normalized Strand
  Cross-Correlation.}
  \label{tab:qc} %% need to make this table
\end{table}

DNA libraries generated by the ChIP-exo protocol seem to be less
complex than the libraries generated by ChIP-Seq, since the possible
number of positions to which the reads can be aligned is being reduced
due to the exonuclease digestion, hence considerable amounts of reads
are being mapped to specific positions. This affects the
interpretation of the PBC, since for ChIP-Seq low PBC values indicate
that the same read has been copied by the amplification process and
aligned multiple times to the same position; while for ChIP-exo when
several reads are aligned to the same position are not necessarily the
same read amplified, but several reads that their $5\prime$ end was
digested to the same position before the amplification step. It is of
special importance to notice that for several ChIP-exo datasets, the
PBC values are quite low. Therefore, by blindly following ChIP-Seq
guidelines those experiment could have been considered as not useful
and repeated.

The Strand Cross-Correlation (SCC) introduced by Kharchenko et al.,
2008 \cite{strandcc} is the most commonly used quality measure in
ChIP-Seq. It is calculated as the correlation between both strand
coverages, where each one is shifted $\delta / 2$ bp towards the
$3\prime$ direction. In general it measures how well the reads mapped
to each strand are clustered around the locations where the proteins
are binding to the DNA, and usually it is expected to observed two
local maxima, one when the profiles are shifted by the average read
length and another when the profiles are shifted by the unobserved
fragment length. In a good ChIP-Seq dataset the last one is also the
SCC global maxima. However, in ChIP-exo's case these two peaks are
confounded. Hence the Normalized Strand Cross-Correlation (NSC) which
is a measure based on the SCC is harder to interpret. Figure
\ref{fig:scc_exo} shows that both local maxima are hard to
differentiate in the SCC curves for the $\sig$ ChIP-exo datasets used
to calculate the NSC values in Table \ref{tab:qc}.

\subsection{Comparison with ChIP-Seq data}

We first compared various factors that could affect binding site
identification between ChIP-exo and ChIP-Seq data. In order to compare
distribution of signal and background between ChIP-exo and ChIP-Seq
data, we calculated ChIP tag counts across the genome by counting the
number of reads mapping to each of 150 non-overlapping
window after extending reads by 150 to their $3\prime$
end directions. ChIP tag counts in ChIP-exo data were linearly related
to ChIP tag counts in ChIP-Seq data for the regions with high ChIP tag
counts (Upper part of Figure \ref{fig:comp}A). This implies that
signals for potential binding sites are well reproducible between
ChIP-exo and ChIP-Seq data. On the other hand, there was clear
difference in the background distribution between them (lower part of
Figure \ref{fig:comp}A). Specifically, in ChIP-Seq data reads were
almost uniformly distributed over background (non-binding) regions and
the ChIP tag counts in there regions were significantly larger than
zero. In contrast, in ChIP-exo data, there was larger variation in
ChIP tag counts among background regions and ChIP tag counts were much
lower in these regions compared to ChIP-Seq data. There were also
large proportion of regions without any read in ChIP-exo data. These
results indicate that for ChIP-exo data a much smaller portion of the
genome is expected to be background.

We next evaluated the ``peak pair'' assumption from Rhee and Pugh,
2011 \cite{exo1}, i.e. a peak of reads in the forward strand is
usually paired with a peak of reads in the reverse strand that is
located in the other site of the binding site. Note that currently
available ChIP-exo data analysis methods, such as Wang et al., 2014
\cite{mace}, Madrigal 2015 \cite{cexor} and Bardet et al., 2013
\cite{peakzilla} rely on this assumption. In order to evaluate this
assumption, we reviewed the proportion of reads in the forward strand
in ChIP-exo peaks such as at least one binding site is predicted in
both ChIP-exo and ChIP-Seq data. We found that strands of reads were
much less balanced in ChIP-exo data than in ChIP-Seq data in these
regions with potential binding sites (Fig. \ref{fig:comp}B) and this
indicates that the peak pair assumption might not hold in real
ChIP-exo data.

We evaluated ChIP-exo data for CTCF factor from human genome
\cite{exo1} to investigate issues specific to eukaryotic genomes for
binding sites identification. Figures \ref{fig:comp}C and
\ref{fig:comp}D display the bin-level average read counts against
mappability and GC content. Each data point is obtained by averaging
the read counts across bins with the same mappability of GC -
content. In Figure \ref{fig:comp}C it is shown that the ChIP-exo tag
counts linearly increases with the mappability score and in Figure
\ref{fig:comp}D it is shown that for GC - content below 0.6, the mean
ChIP tag count increases and for GC - content greater than 0.6 it
shows a decreasing trend. Kuan et al., 2011 \cite{mosaics} studied the
presence of the mappability and GC - content biases in ChIP-Seq's
background. It is not surprising to see these biases also present in
ChIP-exo data, since ChIP-Seq and ChIP-exo signal seems to be linearly
correlated for enriched regions (Figure \ref{fig:comp}A). Rozowsky et
al., 2009 \cite{peakseq} and Valouev et al., 2008 \cite{quest} provide
in depth analysis of the mappability and GC - content biases for
ChIP-Seq respectively. Finally, these results indicate that binding
site identification for ChIP-exo data benefits from using methods that
take into account of apparent sequence biases such as mappability and
GC content.

\subsection{ChIP-exo Quality Control Pipeline}

Figure \ref{fig:qcdiagram} shows a flowchart for the ChIP-exo QC
pipeline. In the first step, we partition the genome by keeping the
non-digested ChIP-exo regions. Then, for each region it counts the
number of fragments that compose the region and the number of
positions to which the reads are being mapped to in each strand. With
these values it calculates the following summary statistics:

\begin{align*}
  \text{ARC} &= \dfrac{\text{Nr. of reads in the region}}{\text{Width of the region}}, \\
  \text{URCR} &= \dfrac{\text{Nr. of reads in the region mapped to
      exactly one position}}{\text{Nr. of reads in the region}}, \\
  \text{FSR} &= \dfrac{\text{Nr. of fwd. strand reads in the
      region}}{\text{Number of reads in the region}}.
\end{align*}

Finally it creates several visualizations designed to diagnose the
quality level of a ChIP-exo sample. Figure \ref{fig:qcdiagram}A shows
the typical behavior of the ARC vs. URCR plot. In general, the plot
depicts two strong arms: One on the left with low ARC values and
varying URCR which corresponds to ChIP-exo's background, regions that
are usually composed by scattered reads that were not digested during
the exonuclease step; and another one where the URCR decreases as the
ARC increases, which corresponds to regions that are usually enriched
and as the URCR decreases the library complexity does it as well, on
the other hand high URCR values correspond to regions composed by
position with relatively few fragments aligned to. Figures
\ref{fig:qcdiagram}B and \ref{fig:qcdiagram}C analyze the strand
imbalance bias in a ChIP-exo experiment: The first one depicts how
quickly the regions exclusively formed by fragments in one strand are
being filtered out as regions with higher depth are observed; and the
second one shows how quickly the FSR's distribution approach the
median, since in a high quality sample it is expected for the median
to be approximately 0.5 and the enriched regions are going to be
composed by fragments sequenced from both strands.

The ChIP-exo QC pipeline provides a model free framework to analyze
the biases in a ChIP-exo experiment by taking advantage of the
exonuclease enzyme that digests the non-enriched regions to partition
the genome, calculates common ChIP-Seq QC metrics in ChIP-exo regions
locally and allows the interpretation of these metrics by the use of
diagnostic plots.

\subsubsection{Enrichment and library complexity in ChIP-exo data}

In ChIP-exo experiments, background fragments are often digested by
the exonuclease enzyme, therefore the balance between the enrichment
and library complexity of an experiment is a key factor determining
the sample's data quality.

Using the Fox A1 in mouse liver cell lines from \cite{exoillumina} and
these two quantities, we explored the relationship between library
complexity and experiment enrichment. In Figure \ref{fig:enrich}A we
present ARC vs. URCR plots for all three replicates. As a case of
study, we compare the three plots to differentiate the quality of the
three experiment. Hence, this might imply that the first replicate to
have have more enriched regions and the third replicate's library
complexity to be lower than the other two replicates library
complexities. To verify this statement, we extracted the sequences
around high confidence binding events and look for the FoxA1 motif
using FIMO \cite{fimo}. Figure \ref{fig:enrich}B shows the number of
candidate regions, which shows that the first replicate is being
allocated into more enriched regions than the other ones. Figure
\ref{fig:enrich}C shows that for the first and third replicates, the
FoxA1 motif is being detected in roughly the same proportion of sites,
and finally in \ref{fig:enrich}D we observe that the first and third
replicate can detect the FoxA1 motif with the same significance, while
the second replicate does not.

\subsubsection{Strand imbalance in ChIP-exo data}

The strand imbalance assessment is based in the observation that the
enriched regions usually are composed of a higher quantity of reads,
therefore we examined the FSR (defined as the ratio of number of
forward stranded reads divided by the total number of reads in a given
region) as the regions with lower depth are being filtered out. This
indicator is of particular importance, as it evaluates the ``peak
pair'' assumption that the original ChIP-exo paper suggested and
multiple ChIP-exo data analysis methods rely on. For every ChIP-exo
experiment, we calculated the global FSR and noticed that for all
experiments is roughly $0.5$, which means there are roughly the same
amount of reads in both strands.

In order to asses the strand imbalance we created the visualization
shown in Figure \ref{fig:strand}: Figure \ref{fig:strand}A presents
the FSR's behavior as the lower depth regions are being filtered out,
while Figure \ref{fig:strand}B) shows which percentage of the regions
are composed by reads in both strand or only one (forward or
backward). In a good data set, it would be expected that all quantiles
shown to be quickly converging towards the median (in panel A) or the
regions composed of reads in one strand being made of few fragments
(in panel B). For each replicate, we divided the partitioned regions
by asking whether they overlap a set of high quality ChIP-exo peaks,
and then we tested (using the Wilcoxon rank sum test over the
imbalance index defined in the methods section) if the strand
imbalance's distribution is the same for both classes. For regions
composed by a higher amount of reads, it is harder to distinguish
their peaks by considering only the strand imbalance, hence in a
better quality ChIP-exo experiment it is easier to distinguish
enriched regions by the amount of reads in both strands. Similarly, we
may consider that the strands of reads for background of a ChIP-exo
experiment is more unbalanced than those for the enriched regions. In
conclusion, Figure \ref{fig:strand} shows that the global FSR does not
represent the experiment's local strand imbalance, hence the ``peak
pair'' assumption does not hold locally in ChIP-exo data.

\subsection{Comparison with ChIP-Seq data using dPeak}

Figure \ref{fig:reso_all} shows comparisons among ChIP-exo, PE
ChIP-Seq and SE ChIP-Seq. We considered the RegulonDB data as ground
truth, since those are the most recent annotation on \emph{Escherichia
  Coli}. A RegulonDB annotation (Salgado et al, 2012 \cite{regulondb})
was considered to be identified if the distance from the closest dPeak
binding site estimate was less that or equal to $20$
bp. That way, the sensitivity is defined as the proportion of
RegulonDB annotations identified in a peak and the resolution is
defined as the minimum distance between a RegulonDB annotation and the
closest dPeak binding site estimate. Figure \ref{fig:reso_all}A shows
that the sensitivity increases as the mean distance between binding
events increases. When the average distance is greater than 200 bp,
dPeak identifies more than the 75\% of the binding events in each
peak, this is intuitive as the mean ChIP-Seq's fragment length is
shorter than 200 bp, hence the read does not contribute to more than
one binding event \cite{dpeak}. When the binding events in a peak are
closer to each other, both ChIP-exo and PE ChIP-Seq are comparable, as
the distance increases ChIP-exo identifies a higher proportion of the
RegulonDB annotations; additionally SE ChIP-Seq is significantly less
sensitive than both ChIP-exo and PE ChIP-Seq. Figure
\ref{fig:reso_all}B shows that ChIP-exo and PE ChIP-Seq are comparable
in resolution, while both protocols significantly outperform SE
ChIP-Seq.

\subsection{Systematic comparison of ChIP-Seq vs ChIP-exo under
  varying sequencing depth}

Previously, ChIP-exo and SE ChIP-Seq have been compared only at a
fixed depth level in the literature, while they did not include PE
ChIP-Seq as well either. In order to address this limitation in
previous studies, we sampled a fixed amount of reads for each of the
ChIP-exo, PE ChIP-Seq and SE ChIP-Seq datasets of the $\sig$ samples
($N$ reads for both ChIP-exo experiment and $N/2$ or $N$ pairs for PE
ChIP-Seq to assume more realistic situation of a fixed cost). For each
sampled dataset we applied our lower-to-higher resolution pipeline by
calling peaks with MOSAiCS \cite{mosaics} and then deconvolving the
binding events by using dPeak \cite{dpeak}. For the ChIP-exo datasets
we called peaks by using the GC-content and mappability models with
MOSAiCS, since it's background is usually composed of scattered reads
across the genome; and for the ChIP-Seq datasets we used their
respective Input samples. Additionally, it is worth noting that for PE
ChIP-Seq we sampled both ends of the fragment, hence for each
sequencing depth we are sampling the half amount of pairs for PE
ChIP-Seq than for ChIP-exo or SE ChIP-Seq.

Figures \ref{fig:design} shows the behavior of each data type in
$\sig$ experiment under aerobic condition when comparable number of
reads is used for all of ChIP-exo, SE ChIP-Seq and PE ChIP-Seq. In
Figure \ref{fig:design}A we show the number of candidate regions
defined as the number of regions where a binding event was identified
in a collection of high quality peaks; in Figure \ref{fig:design}B we
depict the number of binding events; in Figure \ref{fig:design}C we
show the number of identified targets, where a RegulonDB was
considered identified if a binding event was identified in a 15 bp
vicinity of it; and finally Figure \ref{fig:design}D show the
resolution defined as the distance from a RegulonDB annotation to the
closest dPeak prediction. It is remarkable that even when the number
of candidate peaks or the number of predicted events is lower for
ChIP-exo, it outperforms both PE and SE ChIP-Seq in number of
identified targets and resolution.

This may suggest that with ChIP-exo less false positive peaks are
being called and that when the targets are being identified, dPeak
estimates binding locations closer to the true location. Additionally,
we can see that as the read depth increases, all four indicator seem
to stabilize and hit a plateau earlier than the cases for ChIP-Seq,
which may indicate that with ChIP-exo a smaller amount of reads is
needed to identify the same number of targets than ChIP-Seq, but it
may be also possible that this is an artifact occurring due to
ChIP-exo's lower library complexity.

Figure \ref{fig:saturation_rif} shows an analogous analysis but using
the $\sig$ replicates with and without rif treatment. The left, middle
and right columns show the fixed depth against the number of predicted
events, identified targets and resolution being compared at a fixed
depth level. The behavior of this quantities seems to be opposite to
the one as in Figure \ref{fig:design}, hence we used the ChIP-exo QC
pipeline in the fixed depth ChIP-exo experiments. Figure
\ref{fig:exoQC_sat_aero} shows scatter plots of ARC vs URCR for
several fixed sample sizes, as the fixed depth increases the two arm
pattern becomes more distinctive while for lower depth, it seems that
the majority of the sampled reads were aligned to enriched regions. On
the other hand in Figures \ref{fig:exoQC_sat1} to
\ref{fig:exoQC_sat4}, we used the ChIP-exo QC pipeline on the samples
that are outperformed by PE and SE ChIP-Seq. For a fixed low depth, we
can see that the majority of the reads are being aligned to
non-enriched regions since the vertical arm seems to be stronger for
all 2 conditions and replicates; for higher depth we can see low URCR
values being predominant, which indicates that the majority of the
regions being formed by few positions with a higher read
concentration. In low complexity regions, the reads are being aligned
to fewer positions but there is no control over the amount of reads
mapped. Hence, those regions are more likely to being strand-imbalance
which in turn may bias the binding site estimate and therefore
decrease the number of identified targets or increase an experiments
resolution.

% In supplement figures \ref{fig:exoQC_sat1} to \ref{fig:exoQC_sat4}, we
% repeat the same analysis for both replicates under the two rif
% conditions.

\subsection{dPeak outperforms competing methods in discovering closely
  spaced binding events from ChIP-exo and ChIP-Seq data}

Figure \ref{fig:methods_comp} compares the resolution defined as the
minimum distance between a RegulonDB annotation and a binding site
predicted by either Peakzilla \cite{peakzilla}, MACE \cite{mace}, GEM
\cite{gem} or dPeak \cite{dpeak}. In a good dataset such as both of
the ChIP-exo experiments under aerobic (panel 931 and 933) conditions,
all the methods are comparable in resolution, and dPeak slightly
outperforms the rest. On the other hand, for experiments dPeak
identified binding sites with higher resolution than Mace and
Peakzilla, but with low resolution than Gem. This may be due that the
fact that Gem uses sequence information in addition to the aligned
$5\prime$ end counts that dPeak uses. 
 \footnote{For here we may
  probably use only 933 as part of the main article and keep the rest
  for the supplement}

% are comparable if we use the init method with motif but not always
%

\section{Conclusions}
\label{sec:conc}

We made a systematic exploration of several ChIP-exo experiments. We
provided a list of factors that reflect the quality of a ChIP-exo
experiment and we developed a QC pipeline which is capable of
assessing the balance between the enrichment and the library
complexity of a ChIP-exo experiment. Additionally, a set of
diagnostics was established to assess the quality of a ChIP-exo
experiment. While the QC pipeline only requires a set of aligned reads
to give a global overview of a ChIP-exo experiment, this overview
coincides with more elaborate analysis that is computationally more
expensive to perform or requires additional inputs that may not be
available, such as motif detection in a set of high quality regions or
resolution analysis given a set of annotations as gold-standard.

We studied the shared biases between ChIP-exo and ChIP-Seq data, and
noticed that for eukaryotic genomes the relationship between ChIP-Seq
data and either the mappability or the GC content scores are still
present in ChIP-exo. We also examined ChIP-exo's background and
noticed that is significantly different from the ChIP-Seq one, since
it consists of only a small quantity of fragments that was not
digested by the exonuclease enzyme. Additionally, we showed that we
have unbalanced number of reads in forward and reverse strands, and
that in a lower quality ChIP-exo experiment those regions are going to
be harder to differentiate from the possibly enriched regions.

To the extent of our knowledge, we made for the fist time a comparison
between ChIP-exo and PE ChIP-Seq. Using a set of annotations as
gold-standard, we showed that both protocols are comparable in
resolution and that for regions with more than one binding site,
ChIP-exo is more sensitive than both SE and PE ChIP-Seq. We made a
rigorous comparison between fixed depth ChIP-exo, PE ChIP-Seq and SE
ChIP-Seq, and we probed that for sufficiently complex libraries,
ChIP-exo experiments can outperform PE and SE ChIP-Seq in number of
identified targets and resolution. The proposed ChIP-exo QC pipeline
provides a rigorous, easily interpretable, computationally efficient
framework to diagnose if the library complexity of a ChIP-exo
experiment is adequate.

\section{Methods}
\label{sec:methods}

\subsection*{Growth conditions.}

\subsection*{ChIP-exo experiments.}

\subsection*{Definition of current ChIP-Seq QC guidelines.}

The statistics in Table \ref{tab:qc} and the SCC curves from Figure
\ref{fig:scc_exo} were calculated with the \textbf{ChIPUtils} package
version 0.99.0, available in
\url{https://github.com/welch16/ChIPUtils}. This package provides an
easy to use interface to calculate basic quality control metrics and
diagnostic plots for ChIP-Seq data.

\subsubsection*{PCR Bottleneck Coefficient.}

The PCR Bottleneck coefficient is a measure of library complexity in
ChIP-Seq data:

\begin{align}
  \mbox{PBC} = \frac{\text{Nr. of positions to which exactly one
      unique mapping read is aligned}}{\text{Nr. of positions to
      which at least one unique mapping read is aligned}} \nonumber
\end{align}

For human and mouse genome, the ENCODE project states that a PBC value
in the 0 - 0.5 range indicates severe bottlenecking, in the 0.5 - 0.8
range moderate bottlenecking, in the 0.8 - 0.9 range indicates mild
bottlenecking and in the 0.9 - 1 range indicates that there is no
presence of bottlenecking.

\subsubsection*{Strand Cross-Correlation.}

The strand cross-correlation was proposed by Kharchenko et al., 2008
\cite{strandcc} and it may be one of the most used of the ChIP-Seq QC
metrics. The SCC curve is defined as:

\begin{align}
  y(\delta) = \sum_c w_c r\left[ n_c^+ \left(x + \frac{\delta}{2}
    \right), n_c^- \left( x- \frac{\delta}{2} \right)\right],
\label{scc}
\end{align}

where $y(\delta)$ is the SCC for a strand shift $\delta$, $r$ is the
Pearson correlation, $w_c$ is the proportion of reads mapped to
chromosome $c$ and $n_c^S$ is the read count vector for strand $S$ and
chromosome $c$. Additionally, two QC metrics are defined:

\begin{align}
  \mbox{NSC} &= \frac{\max_\delta y(\delta)}{\min_\delta y(\delta)}, \label{nsc} \\
  \mbox{RSC} &= \frac{\max_\delta y(\delta) -
    y_{\text{bgd}}}{y_{\text{rl}} - y_{\text{bgd}}}. \label{rsc}
\end{align}

where $y_{\text{bgd}}$ is the background SCC level and $y_{\text{rl}}$
is the SCC ``phantom peak'' value.

\subsection*{Mappability and GC content scores.}

To define the mappability score we follow the definition from Rozowsky
et al., 2009 \cite{peakseq}:

\begin{align}
  m_i &= \sum_{k = i - L +1}^{i + L - 1} \frac{\delta_k }{2L - 1}.
\end{align}

where $\delta_i$ is the indicator if the base at coordinate $i$ can be
mapped uniquely by a 32 bp sequence at position $i$, and $L$ is the
expected fragment length. GC - content score is defined analogously,
where $\delta_i$ represents the occurrence of a G or C at the $i$-th
position in the genome.

The mappability and GC - content scores for a bin are defined as the
average of the scores across the nucleotides in the bin.

\subsection*{ChIP-exo quality control pipeline.}

We used the R package \textbf{ChIPexoQual} to assess the quality of
the ChIP-exo datasets by following the steps described in Figure
\ref{fig:qcdiagram}. We used version 1.0, and it is available in
\url{https://github.com/welch16/ChIPexoqual}.

\subsection*{Motif analysis of Fox A1 enriched regions}

For all three replicates we called peaks using the MOSAiCS GC +
Mappability model using an FDR level of 5\%, filtering out the peaks
with average ChIP counts below one hundred fragments and merging peaks
gaped by at most 200 bp. Then, we fitted the dPeak model considering
at most 5 binding events for each peak, and we searched for the Fox A1
motif over a 10 bp window around the estimated binding events. We used
FIMO's command line 4.9.1 version \cite{fimo}.

\subsection*{Imbalance index}

For every ChIP-exo experiment, we partitioned the experiment into the
non-digested regions using the QC pipeline. For each region we
calculated the FSR defined as the ratio between the number of forward
stranded reads and the total number of reads in a region.

Then we
transformed the FSR into an imbalance index is defined as:

\[
\mbox{Imbalance index} = -\log_{10} (4 \times \mbox{FSR} \times (1 - \mbox{FSR}))
\]

\subsection*{Construction of a SE ChIP-Seq from a PE ChIP-Seq experiment.}

For the rif-treatment ChIP-Seq experiments, we sampled SE ChIP-Seq
experiment from the PE ones by taking one of both ends randomly
following with equal probability.

\subsection*{dPeak analysis of $\sig$ ChIP-exo and ChIP-Seq data.}

For the resolution and sensitivity analysis, we used the MOSAiCS GC
content + mappability model to call peaks for ChIP-exo experiments,
while for both SE and PE ChIP-Seq experiments we used the MOSAiCS
Input model. To avoid false positives we only considered ChIP-exo
peaks with average ChIP counts greater than 3000 that overlapped both
the SE and PE ChIP-Seq peaks, we considered other cutoff values but
still obtained results similar to what we presented in this
paper. Then we estimated the binding site events using the dPeak model
with a maximum number of 5 binding events. We considered RegulonDB
annotation as gold-standard; Resolution is defined as the minimum
distance from an annotation to an estimated binding events and
Sensitivity is defined as the fraction of annotation within
15 bp from an estimated binding event.

\subsection*{Saturation analysis of ChIP-exo, PE ChIP-Seq and SE ChIP-Seq.}

To perform the saturation analysis, we sub-sampled $N$ fragments for
both ChIP-exo and SE ChIP-Seq protocols. For PE ChIP-Seq we
sub-sampled $N$ pairs or $N/2$ fragments. For each seed, we called
peaks using MOSAiCS \cite{mosaics} (GC content + mappability for
ChIP-exo and Input for SE and PE ChIP-Seq) for the maximum sample size
and to avoid false positives we considered only the top 500 peaks for
each data protocol. We defined the number of candidate regions as the
number of top sample peaks such that a binding events was estimated
using the sampled reads and the dPeak's model; the number of predicted
events is the total quantity of binding events estimated using the
dPeak's model; the number of identified targets are number of
gold-standard annotations within 15 bp from an estimated binding
events; and the resolution is defined as the minimum distance from a
gold-standard annotation to an estimated binding event. We repeated
this analysis for ten seeds and reported the median between all those
values.

\subsection*{Method comparison for ChIP-exo.}

We considered dPeak Chung et al., 2013 \cite{dpeak}, GEM Guo et al.,
2012 \cite{gem}, MACE Wang et al., 2014 \cite{mace} and Peazilla
Bardet et al., 2013 \cite{peakzilla} for the ChIP-exo data
analysis. for the dPeak algorithm we used the R package \textbf{dPeak}
version 2.0.1 which is available from
\url{https://github.com/dongjunchung/dpeak}. For the GEM algorithm, we
used it's Java implementation version 2.6 which is available from
\url{http://groups.csail.mit.edu/cgs/gem/}. For the Mace algorithm, we
used it Python implementation version 1.2, which is available from
\url{http://dldcc-web.brc.bcm.edu/lilab/MACE/docs/html/}. For the
Peakzilla algorithm, we used the version available in
\url{https://github.com/steinmann/peakzilla}. Candidate regions for
\textbf{dPeak} were identified for each replicate of ChIP-exo data
using the \textbf{MOSAiCS} algorithm Kuan et al., 2011 \cite{mosaics}
(one sample analysis using false discovery rate of 0.01\%)
implemented as an R package \textbf{mosaics} version 2.9.7 (available
from \emph{bioconductor}). We further filtered out candidate regions
by using the 300 peaks with higher average ChIP tag count to
avoid potential false positive based on the exploratory
analysis. These regions were also explicitly provided to the GEM
algorithm as candidate regions. Default tuning parameters were used
during model fitting for all methods. We were unable to use CexoR
\cite{cexor} to estimate ChIP-exo binding sites.

\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{chip_exo_paper}
\nocite{exo_gb}
\nocite{maplot1}
\nocite{maplot2}
\nocite{chipbeyond}

% \nocite{esl}

\newpage

\section{Figures}

\begin{figure}[h!]
  \centering
  \includegraphics[width =
  .95 \textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper/chip_explanation_withEnzyme.pdf}
  \caption{\textbf{Description of ChIP-exo, SE ChIP-Seq and PE
      ChIP-Seq:} A TF is bound to the forward (red) and backward
    (blue) strand of the DNA. Then, is sonicated: For ChIP-exo a
    exonuclease enzyme (orange hexagon) trims the $5\prime$ ends of
    each DNA fragment to a fixed distance from the bound protein,
    finally is subjected to Immunoprecipitation and amplification. For
    both ChIP-exo and SE ChIP-Seq an adapter is ligated (green
    triangles) at the $5\prime$ ends, while for PE ChIP-Seq is ligated
    to both ends.}
  \label{fig:chip_diagram}
\end{figure}
\newpage

\begin{figure}[h!]
\centering
\includegraphics[width = .8\textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper/EColi_strand_cross_corr.pdf}
\caption{\textbf{SCC curves for $\sig$ samples.} The ``phantom peak''
  and the summit that corresponds to the read and fragment length
  respectively are confounded due to the exonuclease digestion.}
  \label{fig:scc_exo}
\end{figure}

\newpage

\begin{figure}[h!]
  \centering
  \includegraphics[width = .46\textwidth,page = 3 ]{../figs/for_paper/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
  \includegraphics[width = .46\textwidth]{../figs/for_paper/forward_strand_ratio_comp_old.pdf}
\includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
  \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
  \caption{ A) Hexbin plot of PE ChIP-Seq bin counts vs ChIP-exo bin
    counts. B) Forward Strand Ratio densities for SE ChIP-Seq and
    ChIP-exo peaks. C) Mappability score vs mean ChIP tag counts with
    0.95 confidence bands. D) GC - content score vs mean ChIP tag
    counts with 0.95 confidence bands.}
  \label{fig:comp}
\end{figure}

% ChIP tag counts increase linearly as mappability scores
% increases. D) ChIP tag counts increase linearly as GC content score
% increases when GC content is less than 0.6 and then ChIP tag counts
% decrease as GC content increases. D) Percentage of sites with at
% least p.value.

\newpage

\begin{figure}[h!]
  \centering
  \includegraphics[width = \textwidth]{../figs/for_paper/coverage_diagram3.pdf}
  \caption{\textbf{Flowchart of the ChIP-exo quality control
      pipeline.} A) ARC vs. URCR, this plot presents a global view of
    the balance between library complexity and enrichment. There are
    two arms: First, one with low ARC, which corresponds to regions
    formed by few aligned positions; Second, where the URCR decreases
    as the ARC increases. B) Min. depth vs Proportions of regions,
    this plot shows the strand composition for all the regions formed
    by a min. number of reads. Regions with low depth tend to be
    formed by undigested reads from one unique strand, while regions
    with higher signal are usually formed by reads in both strands. C)
    Min depth vs. FSR, this plot depicts how quickly the FSR's
    distribution approximate the median. In a high quality sample, the
    median is around 0.5, and the other quantiles reach that value
    quickly.}
  \label{fig:qcdiagram}
\end{figure}

\newpage

\begin{figure}[h!]
  \centering
  \includegraphics[width = .9  \textwidth,page = 1]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
  \newline
  \includegraphics[width = .3 \textwidth,page = 1]{../figs/FOXA1_mm9_fimo/FOXA1_summaries_by_replicate.pdf}
  \includegraphics[width = .3 \textwidth,page = 3]{../figs/FOXA1_mm9_fimo/FOXA1_summaries_by_replicate.pdf}
  \includegraphics[width = .3 \textwidth,page = 1]{../figs/FOXA1_mm9_fimo/FOXA1_log10pval_cumulative_nmotifs.pdf}
  \caption{Using the mouse FoxA1 experiment from \cite{exoillumina}:
    A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$, there is a
    slight separation into two strong arms, one corresponds to low
    $\mbox{ARC}$ and varying $\mbox{URCR}$, and for the other
    $\mbox{URCR}$ decreases as $\mbox{ARC}$ increases. B) Number of
    candidate sites for each replicate. C) Percentage of candidate
    sites where the FoxA1 motif was detected. D) Cumulative
    proportion of detected motifs by replicate.}
  \label{fig:enrich}
\end{figure}

\newpage

\begin{figure}[h!]
  \centering  
  \includegraphics[width = .9\textwidth,page = 1]{../figs/Carroll_mice_for_paper/Strand_imbalance.pdf} 
  \newline
  \includegraphics[width = .9\textwidth,page = 2]{../figs/Carroll_mice_for_paper/Strand_imbalance.pdf} 
  \includegraphics[width = .4\textwidth]{../figs/for_paper/Carroll_FSR_depth_VS_pvalWilcoxon.pdf}
  \caption{Strand imbalance QC plots for the same data as in Figure
    \ref{fig:enrich}A. A) FSR distribution quantiles as the lower
    depth regions are being filtered out, all quantiles approach to
    the median as the lower bound increases. B) Stacked histogram with
    the proportion of regions that are formed by two strands or only
    one, in a good sample the single-stranded regions are going to be
    filtered out quickly as in the middle row. C)
    $-\log_{10}(\text{p.value})$ of testing if the imbalance
    distributions differs when ChIP-exo regions overlap their peaks.}
  \label{fig:strand}
\end{figure}

\newpage

\begin{figure}[h!]
  \centering
  \includegraphics[width = .46\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
  \includegraphics[width = .46\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
   \caption{Comparison of (A) sensitivity and (B) resolution between
     ChIP-exo and ChIP-Seq data. Sensitivity is defined as the
     proportion of RegulonDB annotations identified using each
     data. Resolution is defined as the distance between RegulonDB
     annotation and its closest prediction.}
  \label{fig:reso_all}
\end{figure}

\newpage

\begin{figure}[h]
  \centering
  \includegraphics[width = .9\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
  \caption{Comparison of the number of candidate regions (A),
    predicted events (B), identified targets (C) and resolution (D)
    among ChIP-exo, PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations
    are considered as a gold standard. A gold standard binding events
    was deemed identified if a binding event was estimated at a $\pm$
    15 vicinity of it.}
  \label{fig:design}
\end{figure}

% \textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at varying
%   sequencing depths.} 
 \newpage

\begin{figure}[h!]
  \centering
\includegraphics[width = .9\textwidth]{../figs/saturation/K12_alignment/sig70_aerobic_enrichment.pdf}
\caption{Hexbin plots of ARC vs URCR of the $\sig$ ChIP-exo experiment
  under aerobic condition when 10K to 90K reads are being sampled.}
  \label{fig:exoQC_sat_aero}
\end{figure}

\newpage

\begin{figure}[h!]
  \centering
  \includegraphics[width = .9\textwidth]{../figs/for_paper/sig70_methods_comparison_FDR5_topM300.pdf}
  \caption{Comparison of the resolution between dPeak, Gem, Mace and
    Peakzilla methods. Resolution is defined as the minimum distance
    between a RegulonDB annotation and a predicted binding event.}
  \label{fig:methods_comp}
\end{figure}

\newpage

\beginsupplement

\section*{Supplement}
\label{sec:supp}

\emph{Note: I am not sure whether to include several figures on this supplement}

\begin{table}[h!]
%% need to fix format
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{verbatim}
##                      files   nreads       pbc      nsc
## 1: edsn1396_Sig70.sort.bam 13445022 0.9426179 8.865244
## 2: edsn1398_Sig70.sort.bam 16538920 0.9378843 7.031836
## 3: edsn1400_Sig70.sort.bam 11642722 0.8891744 10.77284
## 4: edsn1402_Sig70.sort.bam 16854026 0.9407020 7.936239
## 5: edsn1396_Sig70.sort.bam  6722511 0.6632742  9.01779
## 6: edsn1398_Sig70.sort.bam  8269460 0.5594449 7.179539
## 7: edsn1400_Sig70.sort.bam  5821361 0.6472382 10.89898
## 8: edsn1402_Sig70.sort.bam  8427013 0.5895118 8.124717
\end{verbatim}
\end{kframe}
\end{knitrout}
\caption{Same QC metrics as in table \ref{tab:qc} but applied to
  Landick's chipseq data of the rif experiment}
\end{table}

\begin{figure}[h!]
  \centering % note change to stat_density(geom = "line")
   \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/sigma_delta_old_densities.pdf}
   \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/sigma_delta_old_densities.pdf}   
  \caption{ C) $\delta$ parameter in dPeak measures average distance
    of the reads to their respective binding site. In ChIP-exo data,
    reads were located much closer to the binding site than in SET
    ChIP-Seq. D) $\sigma$ parameter measure the dispersion of reads
    around each binding site. In ChIP-exo data, reads showed less
    variation around the their respective binding sites compared to
    SET ChIP-Seq.}

\end{figure}

\newpage

\begin{figure}[h!]
  \centering
  \includegraphics[width = .7\textwidth]{../figs/saturation/K12_alignment/PBC_saturation.pdf}
  \caption{Average PBC (among all seeds) of the sampled ChIP-exo, PE
    ChIP-Seq and SE ChIP-Seq experiments under the rif treatment
    conditions used for saturation analysis.}
  \label{fig:pbc_saturation}
\end{figure}

\newpage

\begin{figure}[h!] %%% fix this plot as one ggplot, facet_grid(condition ~ variable,scales = "free_y")
  \centering
  \begin{itemize}
  \item Rep-1 and rif-0min:
  \end{itemize}
  \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
  \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
  \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}

  \begin{itemize}
  \item Rep-1 and rif-20min:
  \end{itemize}

  \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
  \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
  \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}

  \begin{itemize}
  \item Rep-2 and rif-0min:
  \end{itemize}

  \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
  \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
  \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}

  \begin{itemize}
  \item Rep-2 and rif-20min:
  \end{itemize}

  \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
  \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
  \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
  \caption{Comparison of the number of predicted events (left),
    identified targets (middle) and resolution (right) among ChIP-exo,
    PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations are considered
    as gold standard. A RegulonDB binding events was deemed identified
    if a binding event was estimated at a $\pm$ 15 vicinity
    of it.}
  \label{fig:saturation_rif}
\end{figure}

% \textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at varying
%   sequencing depth}

\newpage

\begin{figure}[h!]
  \centering
  \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
  \caption{A) ARC vs URCR hexbin plots of Rep-1 and rif-0min from $\sig$ experiment when 100K to 900K reads are being
    sampled for each panel. }
  \label{fig:exoQC_sat1}
\end{figure}

\newpage

\begin{figure}[h!]
  \centering
  \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}

  \caption{ARC vs URCR hexbin plots of Rep-1 and rif-20min from $\sig$ experiment when 100K to 900K reads are being
    sampled for each panel.}
  \label{fig:exoQC_sat2}
\end{figure}

\newpage

\begin{figure}[h!]
  \centering
  \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
  \caption{ARC vs URCR hexbin plots of Rep-2 and rif-0min from $\sig$ experiment when 100K to 900K reads are being
    sampled for each panel.}
  \label{fig:exoQC_sat3}
\end{figure}

\newpage

\begin{figure}[h!]
  \centering
  \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
  \caption{ARC vs URCR hexbin plots of Rep-2 and rif-20min from $\sig$ experiment when 100K to 900K reads are being
    sampled for each panel.}
  \label{fig:exoQC_sat4}
\end{figure}

\end{document}